---
title: "Bail data"
author: "Datathonteam1"
date: "4/24/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(plotly)
library(tibble)
library(scales)
library(sf)
#require(scales)
```

```{r}
defdoc <- readr::read_csv('https://storage.googleapis.com/jat-rladies-2021-datathon/defendant_docket_details.csv')
```

```{r}
baildata <- readr::read_csv('https://storage.googleapis.com/jat-rladies-2021-datathon/bail.csv')
```

```{r}
# Select variables of interest and create new datset
defdoc2 <- select(defdoc, docket_id, gender, race)
```

```{r}
# Select variables of interest and create new datset
baildata2 <- select(baildata, docket_id, action_date, action_type_name, total_amount, participant_name__first_name, participant_name__last_name)
```

```{r}
# turn to tibble 
as_tibble(baildata2)
as_tibble(defdoc2)
```

```{r}
# Create new dataset by merging both datasets
jatdata2 <- merge(defdoc2, baildata2, by="docket_id")
```

```{r}
# Create new variable for judge name 
newjatdata2 <- jatdata2 %>% 
  
  unite(Judge, c(participant_name__first_name, participant_name__last_name), sep = " ", remove = TRUE)
```

```{r}
# Create new variables for year and month
newjatdata2$year <- format(newjatdata2$action_date, format="%Y")

newjatdata2$month <- format(newjatdata2$action_date, format="%m")

# as.numeric to make breaks below
newjatdata2$year <- as.numeric(newjatdata2$year)
newjatdata2$month <- as.numeric(newjatdata2$month)

# Create breaks for 2 year increments and seasonal months
newjatdata2$year <- cut(newjatdata2$year, breaks=c(0, 2005, 2007, 2009, 2011, 2013, 2015, 2017, 2019, 2021), labels=c("2003-2004", "2005-2006", "2007-2008", "2009-2010", "2011-2012", "2013-2014", "2015-2016", "2017-2018", "2019-2020"), right = FALSE)

newjatdata2$month <- cut(newjatdata2$month, breaks=c(0, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13), labels=c("Winter", "Winter", "Spring", "Spring", "Spring", "Summer", "Summer", "Summer", "Fall", "Fall", "Fall", "Winter"), right = FALSE)
```

```{r}
#remove rows with NA
completerecords <- na.omit(newjatdata2)
#na.fail(completerecords)  # trust, but verify!
```

```{r}
# the actual values of the outliers with $out
boxplot(completerecords$total_amount)$out
# Now you can assign the outlier values into a vector
outliers <- boxplot(completerecords$total_amount)$out
```
#more outliers present even after removing. Will not use the removal outliers dataset. 
```{r}
# Now you can remove the rows containing the outliers, one possible option is:

remove_outliers <- completerecords[-which(completerecords$total_amount %in% outliers),]

# If you check now with boxplot, you will notice that those pesky outliers are gone

boxplot(remove_outliers$total_amount)
```

```{r}
completerecords$action_type_name <- as.factor(completerecords$action_type_name)

completerecords$year <- as.factor(completerecords$year)
```


```{r}
is.factor(completerecords$action_type_name)
is.factor(completerecords$year)
```

```{r}
saveRDS(completerecords, file = "new_bail_data.rds")
```


#Judge Marsha Neifield
#facetwrap by both year and action type
```{r}
plot <- completerecords  %>% 
filter(factor(Judge) ==  "Marsha Neifield" )



MN <- ggplot(data = plot, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  geom_col(position="dodge", color="black") +
  facet_wrap(~year + action_type_name) +
  labs(title = "Judge:Marsha Neifield", x = "Season", y = "Bail Amount")

MN + scale_y_log10()

```
#facetwrap by year 
```{r}
MN1 <- ggplot(data = plot, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  geom_col(position="dodge", color="black") +
  facet_wrap(~year) +
  labs(title = "Judge:Marsha Neifield", x = "Season", y = "Bail Amount")

MN1 + scale_y_log10()
```
#facetwrap action type for all years
```{r}
MN2 <- ggplot(data = plot, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  geom_col(position="dodge", color="black") +
   facet_wrap(~action_type_name) +
  labs(title = "Judge:Marsha Neifield", x = "Season", y = "Bail Amount")

MN2 + scale_y_log10()
```
#Filter by specific judge and year then facetwrap by action type
```{r}
plot1 <- completerecords  %>% 
filter(factor(Judge) ==  "Marsha Neifield" & year == "2009-2010") 
  



MN3 <- ggplot(data = plot1, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  geom_col(position="dodge", color="black") +
   facet_wrap(~action_type_name) +
  labs(title = "Judge:Marsha Neifield 2009-2010", x = "Season", y = "Bail Amount")

MN3 + scale_y_log10()

```

#Judge Joseph O'Neill

#use coord_flip() and facetwrap by year 
```{r}
plot2JO <- completerecords  %>%
filter(factor(Judge) ==  "Joseph O'Neill" )

JO1 <- ggplot(data = plot2JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  coord_flip() +
  geom_col(position="dodge", color="black") +
  facet_wrap(~year) +
  labs(title = "Judge:Joseph O'Neill", x = "Season", y = "Bail Amount")

JO1 + scale_y_log10()
```

#use coord_flip() and facetwrap by action type 
```{r}
JO2 <- ggplot(data = plot2JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  coord_flip() +
  geom_col(position="dodge", color="black") +
  facet_wrap(~action_type_name) +
  labs(title = "Judge:Joseph O'Neill", x = "Season", y = "Bail Amount")

JO2 + scale_y_log10()
```

#use coord_flip() and facetwrap by year and action type 
```{r}
JO <- ggplot(data = plot2JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  coord_flip() +
  geom_col(position="dodge", color="black") +
  facet_wrap(~year + action_type_name) +
  labs(title = "Judge:Joseph O'Neill", x = "Season", y = "Bail Amount")

JO + scale_y_log10()
```

#use coord_flip() and filter by specific judge and year then facetwrap by action type
```{r}
plot3JO <- completerecords  %>% 
filter(factor(Judge) ==  "Joseph O'Neill" & year == "2009-2010") 
  


JO3 <- ggplot(data = plot3JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
   coord_flip() +
  geom_col(position="dodge", color="black") +
   facet_wrap(~action_type_name) +
  theme(legend.position="bottom") +
  labs(title = "Judge: Joseph O'Neill 2009-2010", x = "Season", y = "Bail Amount")

JO3 + scale_y_log10()

```


New Section-1

#use coord_flip() and filter by specific judge and year then facetwrap by action type and use scale limit upto 1 million dollars instead of scale log10
```{r}
JO5 <- ggplot(data = plot3JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
   coord_flip() +
   geom_col(position="dodge", color="black") +
   facet_wrap(~action_type_name, scales = "free") +
  theme(legend.position="bottom") +
  labs(title = "Judge:Joseph O'Neill 2009-2010", x = "Season", y = "Bail Amount")

ggplotly(JO5 + scale_y_continuous(limits = c(0, 1000000)), tooltip = c("y", "x")) %>%
  layout(legend = list(orientation = 'h', x=0.1, xanchor="bottom", yanchor="center", y=-0.3))

```


```{r}
JO4 <- ggplot(data = plot3JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
   coord_flip() +
  geom_point(pch = 21, position = position_jitterdodge()) +
   facet_wrap(~action_type_name, scales = "free") +
  theme(legend.position="bottom") +
  labs(title = "Judge:Joseph O'Neill 2009-2010", x = "Season", y = "Bail Amount")

ggplotly(JO4 + scale_y_continuous(limits = c(0, 1000000)), tooltip = c("y", "x")) %>%
  layout(legend = list(orientation = 'h', x=0.1, xanchor="bottom", yanchor="center", y=-0.3)) 
```

New Section-2

#Grouping by race & action type then taking mean bail amount for each judge based off those groupings. Plot this information using ggplot. Example code below. 
```{r}
df <- completerecords %>%
  group_by(race, action_type_name) %>%
  filter(factor(Judge) ==  "Joseph O'Neill") %>%
  summarize(mean_total_amount =mean(total_amount)) %>%
  mutate(race=fct_reorder(race, mean_total_amount)) 
df

  
```

# Plot the df object with the mean of bail amount per race and action type for judge Joseph O'Neill and scale for max amount in table above. No mean was over $150,000 dollars so plotted for max $150,000 dollars. 
```{r}
JO7 <- ggplot(data = df, aes(x = factor(action_type_name), y = mean_total_amount, fill = race )) +
   coord_flip() +
   geom_col(position="dodge", color="black") +
  labs(title = "Judge: Joseph O'Neill", subtitle = "Municipal Court", caption = "Mean bail amount for all years judge has been in office grouped by race and action type.", x = "Action Type", y = "Mean Bail Amount")

 ggplotly(JO7 + scale_y_continuous(limits = c(0, 150000)))
```

New Section-3 

#Fix the spacing for the x-axis by repositioning the legend to the bottom using theme. NOTE:Use theme to change the way plot looks. 

```{r}

JO6 <- ggplot(data = plot3JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
   coord_flip() +
   #geom_col(position="dodge", color="black") +
  geom_point(pch = 21, position = position_jitterdodge()) +
   facet_wrap(~action_type_name, scales = "free") +
  theme(legend.position="bottom") +
  labs(title = "Judge:Joseph O'Neill 2009-2010", caption = "Viewing bail amount upto 1 million dollars only", subtitle = "Track record for 2 years (2009-2010)", x = "Season", y = "Bail Amount")

JO6 + scale_y_continuous(limits = c(0, 1000000))
```

New Section-4 

#Plot count of action type per judged grouped by race
```{r}
df2 <- completerecords %>% 
  group_by(race, action_type_name) %>%
  filter(factor(Judge) ==  "Rayford Means") %>% 
  summarise(counts = n()) %>%
  mutate(race=fct_reorder(race, counts))
df2
```

```{r}
JO8 <- ggplot(data = df2, aes(x = action_type_name, y = counts, fill = race )) +
   coord_flip() +
   geom_col(position="dodge", color="black") +
  labs(title = "Judge: Rayford Means", subtitle = "Municipal Court", caption = "Sum of all types of bail actions taken for all years judge has been in office grouped by race.", x = "Action Type", y = "Count")

 ggplotly(JO8 + scale_y_continuous(limits = c(0, 15000)))
```

#Plot count of action type per judged grouped by gender
```{r}
df3 <- completerecords %>% 
  group_by(gender, action_type_name) %>%
  filter(factor(Judge) ==  "Rayford Means") %>% 
  summarise(counts = n()) %>%
  mutate(gender=fct_reorder(gender, counts))
df3
```


```{r}
JO9 <- ggplot(data = df3, aes(x = action_type_name, y = counts, fill = gender)) +
   coord_flip() +
   geom_col(position="dodge", color="black") +
  labs(title = "Judge: Rayford Means", subtitle = "Municipal Court", caption = "Sum of all types of bail actions taken for all years judge has been in office grouped by race.", x = "Action Type", y = "Count")

 ggplotly(JO9 + scale_y_continuous(limits = c(0, 15000)))
```















