---
title: "Bail data"
author: "Datathonteam1"
date: "4/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(tibble)
require(scales)
```

```{r}
defdoc <- readr::read_csv('https://storage.googleapis.com/jat-rladies-2021-datathon/defendant_docket_details.csv')
```

```{r}
baildata <- readr::read_csv('https://storage.googleapis.com/jat-rladies-2021-datathon/bail.csv')
```

```{r}
# Select variables of interest and create new datset
defdoc2 <- select(defdoc, docket_id, gender, race)
```

```{r}
# Select variables of interest and create new datset
baildata2 <- select(baildata, docket_id, action_date, action_type_name, total_amount, participant_name__first_name, participant_name__last_name)
```

```{r}
# turn to tibble 
as_tibble(baildata2)
as_tibble(defdoc2)
```

```{r}
# Create new dataset by merging both datasets
jatdata2 <- merge(defdoc2, baildata2, by="docket_id")
```

```{r}
# Create new variable for judge name 
newjatdata2 <- jatdata2 %>% 
  
  unite(Judge, c(participant_name__first_name, participant_name__last_name), sep = " ", remove = TRUE)
```

```{r}
# Create new variables for year and month
newjatdata2$year <- format(newjatdata2$action_date, format="%Y")

newjatdata2$month <- format(newjatdata2$action_date, format="%m")

# as.numeric to make breaks below
newjatdata2$year <- as.numeric(newjatdata2$year)
newjatdata2$month <- as.numeric(newjatdata2$month)

# Create breaks for 2 year increments and seasonal months
newjatdata2$year <- cut(newjatdata2$year, breaks=c(0, 2005, 2007, 2009, 2011, 2013, 2015, 2017, 2019, 2021), labels=c("2003-2004", "2005-2006", "2007-2008", "2009-2010", "2011-2012", "2013-2014", "2015-2016", "2017-2018", "2019-2020"), right = FALSE)

newjatdata2$month <- cut(newjatdata2$month, breaks=c(0, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13), labels=c("Winter", "Winter", "Spring", "Spring", "Spring", "Summer", "Summer", "Summer", "Fall", "Fall", "Fall", "Winter"), right = FALSE)
```

```{r}
#remove rows with NA
completerecords <- na.omit(newjatdata2)
```

```{r}
# the actual values of the outliers with $out
boxplot(completerecords$total_amount)$out
# Now you can assign the outlier values into a vector
outliers <- boxplot(completerecords$total_amount)$out
```
#more outliers present even after removing. Will not use the removal outliers dataset. 
```{r}
# Now you can remove the rows containing the outliers, one possible option is:

remove_outliers <- completerecords[-which(completerecords$total_amount %in% outliers),]

# If you check now with boxplot, you will notice that those pesky outliers are gone

boxplot(remove_outliers$total_amount)
```

```{r}
completerecords$action_type_name <- as.factor(completerecords$action_type_name)

completerecords$year <- as.factor(completerecords$year)
```

```{r}
is.factor(completerecords$action_type_name)
is.factor(completerecords$year)
```

```{r}
saveRDS(completerecords, file = "new_bail_data.rds")
```


#Judge Marsha Neifield
#facetwrap by both year and action type
```{r}
plot <- completerecords  %>% 
filter(factor(Judge) ==  "Marsha Neifield" )



MN <- ggplot(data = plot, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  geom_col(position="dodge", color="black") +
  facet_wrap(~year + action_type_name) +
  labs(title = "Judge:Marsha Neifield", x = "Season", y = "Bail Amount")

MN + scale_y_log10()

```
#facetwrap by year 
```{r}
MN1 <- ggplot(data = plot, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  geom_col(position="dodge", color="black") +
  facet_wrap(~year) +
  labs(title = "Judge:Marsha Neifield", x = "Season", y = "Bail Amount")

MN1 + scale_y_log10()
```
#facetwrap action type for all years
```{r}
MN2 <- ggplot(data = plot, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  geom_col(position="dodge", color="black") +
   facet_wrap(~action_type_name) +
  labs(title = "Judge:Marsha Neifield", x = "Season", y = "Bail Amount")

MN2 + scale_y_log10()
```
#Filter by specific judge and year then facetwrap by action type
```{r}
plot1 <- completerecords  %>% 
filter(factor(Judge) ==  "Marsha Neifield" & year == "2009-2010") 
  



MN3 <- ggplot(data = plot1, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  geom_col(position="dodge", color="black") +
   facet_wrap(~action_type_name) +
  labs(title = "Judge:Marsha Neifield 2009-2010", x = "Season", y = "Bail Amount")

MN3 + scale_y_log10()

```

#Judge Joseph O'Neill

#use coord_flip() and facetwrap by year 
```{r}
plot2JO <- completerecords  %>%
filter(factor(Judge) ==  "Joseph O'Neill" )

JO1 <- ggplot(data = plot2JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  coord_flip() +
  geom_col(position="dodge", color="black") +
  facet_wrap(~year) +
  labs(title = "Judge:Joseph O'Neill", x = "Season", y = "Bail Amount")

JO1 + scale_y_log10()
```

#use coord_flip() and facetwrap by action type 
```{r}
JO2 <- ggplot(data = plot2JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  coord_flip() +
  geom_col(position="dodge", color="black") +
  facet_wrap(~action_type_name) +
  labs(title = "Judge:Joseph O'Neill", x = "Season", y = "Bail Amount")

JO2 + scale_y_log10()
```

#use coord_flip() and facetwrap by year and action type 
```{r}
JO <- ggplot(data = plot2JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
  coord_flip() +
  geom_col(position="dodge", color="black") +
  facet_wrap(~year + action_type_name) +
  labs(title = "Judge:Joseph O'Neill", x = "Season", y = "Bail Amount")

JO + scale_y_log10()
```

#use coord_flip() and filter by specific judge and year then facetwrap by action type
```{r}
plot3JO <- completerecords  %>% 
filter(factor(Judge) ==  "Joseph O'Neill" & year == "2009-2010") 
  


JO3 <- ggplot(data = plot3JO, aes(x = factor(month), y = total_amount, fill = factor(race))) +
   coord_flip() +
  geom_col(position="dodge", color="black") +
   facet_wrap(~action_type_name) +
  labs(title = "Judge:Joseph O'Neill 2009-2010", x = "Season", y = "Bail Amount")

JO3 + scale_y_log10()

```









