---
title: "Philadelphia Judges"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(ggplot2)
library(dplyr)

# I performed most data cleaning separately and saved as Rds 
# I could do a better job here and also do this separately

# This is offenses & dispositions data that has been cleaned up
od_clean <- readRDS('~/Documents/od_clean.Rds')%>% 
  rename(judge = disposing_authority__document_name) %>% 
  mutate(grade_desc = paste(grade,description_clean,sep="_")) %>% 
  filter(!is.na(disposition))

# This is defendant and docket info table merged with defendant IDs
ddd <- readRDS('~/Documents/ddd.Rds')

# Again, could do this outside app
merged <- od_clean %>% 
  left_join(ddd, by = "docket_id") %>% 
  mutate(year = lubridate::year(filing_date)) %>% 
  # We dont end up using most the data in the current app
  select(judge, year, docket_id, grade, description, description_clean,
         gender, defendant_id, race, grade_desc, sentence_type)
# Display options in order of most common
judge_options <- na.omit(dplyr::pull(dplyr::count(od_clean, judge, sort=T), judge))
description_options <- na.omit(dplyr::pull(dplyr::count(od_clean, description_clean, sort=T), description_clean))
  
```


Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}

# Now create the inputs
sliderInput('year', 'Year Range (filing date)', min=2010, max=2020,
            value=c(2010, 2020), step=1, round=T, sep="")
selectInput('judge', 'Judge', judge_options)
selectInput('grade', "Grade", unique(od_clean$grade), 
            selected = unique(od_clean$grade),
            multiple = TRUE)
# For now, just pick one Offense
selectInput("description", "Offense Description", description_options, 
            selected = description_options[1:3],
            multiple = FALSE)
# Alternatively, could have free text input here and use grep to find offenses associated
# textInput('search_term','Offense Search Term')
selectInput('defendant_gender', 'Defendant Gender', choices = c("Male","Female"),
            multiple = TRUE, selected = c("Male","Female"))


```

The options shown are in the order of the most common judges and descriptions in the dataset.

```{r}
# Modify the data based on these searches
data_input <- reactive({
  merged %>% 
  mutate(select_judge = ifelse(judge==input$judge, T, F)) %>% 
  filter(gender %in% input$defendant_gender,
         # grepl(pattern = tolower(input$search_term), unlist(tolower(description_clean))),
         grade %in% input$grade,
         description_clean %in% input$description,
         year >= input$year[1] & year <= input$year[2])
})

data_judge_filter <- reactive({
  merged %>% 
  mutate(select_judge = ifelse(judge==input$judge, T, F)) %>% 
  filter(judge==input$judge)
})

```


Column {.tabset}
-----------------------------------------------------------------------

### Number of dockets

Results filtered based on the options selected.
```{r}
renderPlot({
  ggplot(data_input(), aes(x=year, fill=select_judge)) + 
  geom_bar(position = "dodge") + 
  labs(title="Dockets per year",
       subtitle = paste0("Selected judge: ", input$judge)) + 
  scale_fill_manual(values = c("lightgray","goldenrod3")) + 
  theme_minimal()

})
```

### Most common offenses for selected judge

These are all dockets, not just those selected on the left. This can help select common offenses adjudicated by the selected judge.
```{r}
renderPlot({
  data_judge_filter() %>% 
    filter(select_judge) %>% 
    mutate(description_clean = forcats::fct_lump(description_clean, n= 5)) %>% 
  ggplot(aes(x=description_clean)) + 
  geom_bar(position = "dodge") + 
  labs(title="Offense descriptions",
       subtitle = paste0("Selected judge: ", input$judge)) + 
  scale_fill_manual(values = c("lightgray","goldenrod3")) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle=90, hjust=1,vjust=0.5))

})
```

Column {.tabset}
-----------------------------------------------------------------------

### Sentence type

```{r}

renderPlot({
  data_input() %>% 
  group_by(grade_desc, select_judge) %>% 
  count(sentence_type) %>% 
  mutate(prop_sentence_type = n/sum(n)) %>% 
  ungroup() %>% 
  mutate(axis_label = paste0(sentence_type, " (n=", as.character(n), ")")) %>% 
    ggplot(aes(x=sentence_type, 
               y = 100*prop_sentence_type, 
               fill=select_judge)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_text(aes(label=n), size=2,
            position=position_dodge(width=0.9), vjust=-0.25) + 
  facet_grid(.~grade_desc, scales = "free_x",space="free_x",
             labeller = labeller(grade_desc = label_wrap_gen(25)))+
  labs(title="Sentence type by judge and offense",
       subtitle = "Selected judge",
       x = "", y = "Percentage",
       caption = "Only considering offenses with a disposition") + 
  scale_fill_manual(values = c("lightgray","goldenrod3")) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle=90, hjust=1,vjust=0.5))
})
```

### Defendant Race

Data shown is condensed to a single defendant per docket (i.e. if there are multiple occurences of the selected offense for a given defendant, we only count the defendant once)
```{r}
renderPlot({
  data_input() %>% 
    distinct(defendant_id, .keep_all = TRUE) %>% 
  ggplot(aes(fill=race,x=select_judge)) + 
  geom_bar(position = "fill") + 
  labs(title=input$judge, y = "Percentage by race") + 
  # scale_fill_manual(values = c("lightgray","goldenrod3")) + 
    scale_y_continuous(labels = scales::percent) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle=90, hjust=1,vjust=0.5))
})
```

Other notes.

### Defendant Gender

Data shown is condensed to a single defendant per docket (i.e. if there are multiple occurences of the selected offense for a given defendant, we only count the defendant once)
```{r}
renderPlot({
  data_input() %>% 
    distinct(defendant_id, .keep_all = TRUE) %>% 
  ggplot(aes(fill=gender,x=select_judge)) + 
  geom_bar(position = "fill") + 
  labs(title=input$judge, y = "Percentage by gender") + 
  scale_fill_manual(values = c("red","blue")) + 
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
})
```

Other notes or table could go here.