---
title: "Dummy Variable Models"
author: "Shanti Agung"
date: "5/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This notebook is an attempt to model judge harshness using dummy variable models.

```{r}
library(patchwork)
library(tidyverse)
library(lme4)
library(tweedie)
library(statmod)
```

```{r}
# Source the functions
source(file.path(rprojroot::find_root(rprojroot::is_rstudio_project),
                 "analyses", "team2", "preprocess", "sa_functions.R"))
```

```{r}
# Make sure the source file (docket details data) exists
source_file <- file.path(rprojroot::find_root(rprojroot::is_rstudio_project),
                         "data", "docket_details.csv")
stopifnot(file.exists(source_file))
```

# Load dataset

```{r}
# load dataset
ddcols = cols(
  docket_id = col_double(),
  gender = col_character(),
  race = col_character(),
  date_of_birth = col_date(format = ""),
  arrest_date = col_date(format = ""),
  complaint_date = col_date(format = ""),
  disposition_date = col_date(format = ""),
  filing_date = col_date(format = ""),
  initiation_date = col_date(format = ""),
  status_name = col_character(),
  court_office__court__display_name = col_character(),
  current_processing_status__processing_status = col_character(),
  current_processing_status__status_change_datetime = col_date(format = ""),
  municipality__name = col_character(),
  municipality__county__name = col_character(),
  judicial_districts = col_character(),
  court_office_types = col_character(),
  court_types = col_character(),
  representation_type = col_character(),
  M = col_double(),
  M3 = col_double(),
  F1 = col_double(),
  F3 = col_double(),
  F2 = col_double(),
  M1 = col_double(),
  M2 = col_double(),
  S = col_double(),
  `F` = col_double(),
  IC = col_double(),
  H2 = col_double(),
  H1 = col_double(),
  S2 = col_double(),
  S1 = col_double(),
  S3 = col_double(),
  judge_id = col_double(),
  disposing_authority__first_name = col_character(),
  disposing_authority__middle_name = col_character(),
  disposing_authority__last_name = col_character(),
  number_prior_dockets = col_double(),
  total_confinement_days = col_double(),
  max_confinement_days = col_double(),
  age = col_double(),
  court_types_cp = col_double(),
  court_types_mc = col_double(),
  court_types_pac = col_double(),
  court_office_types_commonwealth = col_double(),
  court_office_types_criminal = col_double(),
  court_office_types_municipal = col_double(),
  court_office_types_supreme = col_double(),
  court_office_types_suprerior = col_double()
)

dockets <- readr::read_csv(source_file, col_types = ddcols)
```

```{r}
# Subset data for analysis
dockets_data <- subset_dockets(dockets)
```

```{r}
glimpse(dockets_data)
```


```{r}
# Create factors
gender_levels <- c("Female", "Male")
race_levels <- c("Asian", "Asian/Pacific Islander", "Bi-Racial", "Black",
                 "Native American/Alaskan Native", "Unknown/Unreported",
                 "White")

dockets_data <- dockets_data %>% 
  mutate(gender = factor(gender, level = gender_levels),
         race = factor(race, level = race_levels),
         court_types_cp = factor(court_types_cp),
         court_types_mc = factor(court_types_mc),
         court_types_pac = factor(court_types_pac),
         judge_id = factor(judge_id),
         year = factor(year, ordered = TRUE))
```

```{r}
# Log transform covariates
dockets_data <- dockets_data %>% 
  mutate(log_age = log(age),
         log_prior_dockets = log(number_prior_dockets + 0.5),
         log_M = log(M + 0.5),
         log_M1 = log(M1 + 0.5),
         log_M2 = log(M2 + 0.5),
         log_M3 = log(M3 + 0.5),
         log_F = log(`F` + 0.5),
         log_F1 = log(F1 + 0.5),
         log_F2 = log(F2 + 0.5),
         log_F3 = log(F3 + 0.5),
         log_S = log(S + 0.5),
         log_S1 = log(S1 + 0.5),
         log_S2 = log(S2 + 0.5),
         log_S3 = log(S3 + 0.5),
         log_IC = log(IC + 0.5),
         log_H1 = log(H1 + 0.5),
         log_H2 = log(H2 + 0.5)
         )
```

Build a random sample.
```{r}
set.seed(30318)
dockets_sample <- sample_n(dockets_data, size = 1000)
```

```{r}
glimpse(dockets_sample)
```


# Tweedie GLM

## Dependent variable: total confinement days

### on docket samples: test a small model, without grade severity

Estimate $\xi$:
```{r}
judge_total_profile_m1 <- tweedie.profile(total_confinement_days ~ log_prior_dockets +
                                            age + gender + race + year + month + judge_id,
                                          do.plot = TRUE, data = dockets_sample)
```

```{r}
xi_est_m1 <- judge_total_profile_m1$xi.max
xi_est_m1
```


```{r}
judge_total_m1 <- glm(total_confinement_days ~ log_prior_dockets + age + gender + race +
                        year + month + judge_id, data = dockets_sample,
                      family = tweedie(var.power = xi_est_m1,
                                       link.power = 0))
```

Diagnostics
```{r}
qqnorm(qresid(judge_total_m1), main = "Quantile residuals (set 1)", las=1); qqline(qresid(judge_total_m1))
```

```{r}
qqnorm(qresid(judge_total_m1), main = "Quantile residuals (set 2)", las=1); qqline(qresid(judge_total_m1))
```
### on docket samples: model that includes grade severity

Estimate $\xi$:
```{r}
judge_total_profile_m2 <- tweedie.profile(total_confinement_days ~ log_prior_dockets +
                                            age + gender + race + year + month + judge_id +
                                            log_M + log_M1 + log_M2 + log_M3 +
                                            log_F + log_F1 + log_F2 + log_F3 +
                                            log_S + log_IC +
                                            log_H1 + log_H2 +
                                            court_types_cp + court_types_mc + court_types_pac,
                                          do.plot = TRUE, data = dockets_sample)
```
```{r}
xi_est_m2 <- judge_total_profile_m2$xi.max
xi_est_m2
```

```{r}
judge_total_m2 <- glm(total_confinement_days ~ log_prior_dockets +
                                            age + gender + race + year + month + judge_id +
                                            log_M + log_M1 + log_M2 + log_M3 +
                                            log_F + log_F1 + log_F2 + log_F3 +
                                            log_S + log_IC +
                                            log_H1 + log_H2 +
                                            court_types_cp + court_types_mc + court_types_pac,
                      data = dockets_sample,
                      family = tweedie(var.power = xi_est_m2, link.power = 0))
```


Diagnostics
```{r}
qqnorm(qresid(judge_total_m2), main = "Quantile residuals (set 1)", las=1); qqline(qresid(judge_total_m2))

```
```{r}
qqnorm(qresid(judge_total_m2), main = "Quantile residuals (set 2)", las=1); qqline(qresid(judge_total_m2))
```
### on full dockets: model that includes log grade severity


Estimate $\xi$:
```{r}
judge_total_profile_m3 <- tweedie.profile(total_confinement_days ~ log_prior_dockets +
                                            age + gender + race + year + month + judge_id +
                                            log_M + log_M1 + log_M2 + log_M3 +
                                            log_F + log_F1 + log_F2 + log_F3 +
                                            log_S + log_IC +
                                            log_H1 + log_H2 +
                                            court_types_cp + court_types_mc + court_types_pac,
                                          do.plot = TRUE, data = dockets_data)
```
```{r}
#xi_est_m3 <- judge_total_profile_m3$xi.max
xi_est_m3 <- 1.463265
xi_est_m3
```

xi_est_m3 = 1.463265

```{r}
judge_total_m3 <- glm(total_confinement_days ~ log_prior_dockets +
                                            age + judge_id + gender + race + year + month + 
                                            log_M + log_M1 + log_M2 + log_M3 +
                                            log_F + log_F1 + log_F2 + log_F3 +
                                            log_S + log_IC +
                                            log_H1 + log_H2 +
                                            court_types_cp + court_types_mc + court_types_pac - 1,
                      data = dockets_data,
                      family = tweedie(var.power = xi_est_m3, link.power = 0))
```

```{r}
summary(judge_total_m3)
```


#### Diagnostics



```{r}
qresid1 <- qresid(judge_total_m3)   # Quantile resids, replication 1
qresid2 <- qresid(judge_total_m3)   # Quantile resids, replication 2
```

```{r}
qqnorm(qresid1, main = "Quantile residuals (set 1)", las=1); qqline(qresid1)
```
```{r}
qqnorm(qresid2, main = "Quantile residuals (set 2)", las=1); qqline(qresid2)
```

```{r}
p1 <- data.frame(qresid = qresid1, mu_hat = fitted(judge_total_m3)) %>% 
  ggplot(aes(x = mu_hat, y = qresid)) +
  geom_point(alpha = 0.5, color = "aquamarine3", size = 2) +
  geom_smooth(color = "aquamarine4", se = FALSE) +
  geom_hline(yintercept = 0) +
  labs(x = "Fitted values",
       y = "Quantile residuals")
```

```{r}
p1
```
```{r}
plot(qresid2 ~ fitted(judge_total_m3), las=1,
     xlab = "Fitted values", ylab = "Quantile residuals")
```
```{r}
p2 <- data.frame(qresid = qresid1, log_prior_dockets = dockets_data$log_prior_dockets) %>% 
  ggplot(aes(x = log_prior_dockets, y = qresid)) +
  geom_point(alpha = 0.5, color = "aquamarine3", size = 2) +
  geom_smooth(color = "aquamarine4", se = FALSE) +
  labs(x = "Log(number of prior dockets)",
       y = "Quantile residuals")
```

```{r}
p2
```
```{r}
plot_qresid_explanatory <- function(explanatory_var, x_label){
 tmp <- data.frame(qresid = qresid1, explanatory_var) %>% 
  ggplot(aes(x = explanatory_var, y = qresid)) +
  geom_point(alpha = 0.5, color = "aquamarine3", size = 2) +
  geom_smooth(color = "aquamarine4", se = FALSE) +
  labs(x = x_label,
       y = "Quantile residuals")
 return(tmp)
}
```

```{r}
p3 <- plot_qresid_explanatory(dockets_data$age, "Age")
p3
```

```{r}
p4 <- plot_qresid_explanatory(dockets_data$gender, "Gender")
p4
```
```{r}
p5 <- plot_qresid_explanatory(dockets_data$race, "Race") +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
p5
```
```{r}
p6 <- plot_qresid_explanatory(dockets_data$judge_id, "Judge ID") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 3))
p6
```
```{r}
p7 <- plot_qresid_explanatory(dockets_data$log_M, "Log(number of grade M offenses)")
p7
```

```{r}
p8 <- plot_qresid_explanatory(dockets_data$log_M1, "Log(number of grade M1 offenses)")
p8
```
```{r}
p9 <- plot_qresid_explanatory(dockets_data$log_M2, "Log(number of grade M2 offenses)")
p9
```
```{r}
p10 <- plot_qresid_explanatory(dockets_data$log_M3, "Log(number of grade M3 offenses)")
p10
```
```{r}
p11 <- plot_qresid_explanatory(dockets_data$log_F, "Log(number of grade F offenses)")
p11
```

```{r}
p12 <- plot_qresid_explanatory(dockets_data$log_F1, "Log(number of grade F1 offenses)")
p12
```
```{r}
p13 <- plot_qresid_explanatory(dockets_data$log_F2, "Log(number of grade F2 offenses)")
p13
```
```{r}
p14 <- plot_qresid_explanatory(dockets_data$log_F3, "Log(number of grade F3 offenses)")
p14
```
```{r}
p15 <- plot_qresid_explanatory(dockets_data$log_S, "Log(number of grade S offenses)")
p15
```

```{r}
p16 <- plot_qresid_explanatory(dockets_data$log_IC, "Log(number of grade IC offenses)")
p16
```

```{r}
p17 <- plot_qresid_explanatory(dockets_data$log_H1, "Log(number of grade H1 offenses)")
p17
```
```{r}
p18 <- plot_qresid_explanatory(dockets_data$log_H2, "Log(number of grade H2 offenses)")
p18
```
```{r}
plot(cooks.distance(judge_total_m3), type = "h", las=1,
     ylab = "Cook's distance, D")
```

```{r}
total_m3_inf <- influence.measures(judge_total_m3)
colSums(total_m3_inf$is.inf)
```


