---
title: "Random Coefficient Models"
author: "Shanti Agung"
date: "4/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook is an attempt to model judge harshness using random coefficient (mixed effects) model.

```{r}
library(patchwork)
library(tidyverse)
library(lme4)
```

```{r}
# Source the functions
source(file.path(rprojroot::find_root(rprojroot::is_rstudio_project),
                 "analyses", "team2", "preprocess", "sa_functions.R"))
```

```{r}
# Make sure the source file (docket details data) exists
source_file <- file.path(rprojroot::find_root(rprojroot::is_rstudio_project),
                         "data", "docket_details.csv")
stopifnot(file.exists(source_file))
```

# Load dataset

```{r}
# load dataset
ddcols = cols(
  docket_id = col_double(),
  gender = col_character(),
  race = col_character(),
  date_of_birth = col_date(format = ""),
  arrest_date = col_date(format = ""),
  complaint_date = col_date(format = ""),
  disposition_date = col_date(format = ""),
  filing_date = col_date(format = ""),
  initiation_date = col_date(format = ""),
  status_name = col_character(),
  court_office__court__display_name = col_character(),
  current_processing_status__processing_status = col_character(),
  current_processing_status__status_change_datetime = col_date(format = ""),
  municipality__name = col_character(),
  municipality__county__name = col_character(),
  judicial_districts = col_character(),
  court_office_types = col_character(),
  court_types = col_character(),
  representation_type = col_character(),
  M = col_double(),
  M3 = col_double(),
  F1 = col_double(),
  F3 = col_double(),
  F2 = col_double(),
  M1 = col_double(),
  M2 = col_double(),
  S = col_double(),
  `F` = col_double(),
  IC = col_double(),
  H2 = col_double(),
  H1 = col_double(),
  S2 = col_double(),
  S1 = col_double(),
  S3 = col_double(),
  judge_id = col_double(),
  disposing_authority__first_name = col_character(),
  disposing_authority__middle_name = col_character(),
  disposing_authority__last_name = col_character(),
  number_prior_dockets = col_double(),
  total_confinement_days = col_double(),
  max_confinement_days = col_double(),
  age = col_double(),
  court_types_cp = col_double(),
  court_types_mc = col_double(),
  court_types_pac = col_double(),
  court_office_types_commonwealth = col_double(),
  court_office_types_criminal = col_double(),
  court_office_types_municipal = col_double(),
  court_office_types_supreme = col_double(),
  court_office_types_suprerior = col_double()
)

dockets <- readr::read_csv(source_file, col_types = ddcols)
```

```{r}
# Subset data for analysis
dockets_data <- subset_dockets(dockets)
```

```{r}
glimpse(dockets_data)
```

# Preprocess data

```{r}
# Create factors
gender_levels <- c("Female", "Male")
race_levels <- c("Asian", "Asian/Pacific Islander", "Bi-Racial", "Black",
                 "Native American/Alaskan Native", "Unknown/Unreported",
                 "White")

dockets_data <- dockets_data %>% 
  mutate(gender = factor(gender, level = gender_levels),
         race = factor(race, level = race_levels),
         court_types_cp = factor(court_types_cp),
         court_types_mc = factor(court_types_mc),
         court_types_pac = factor(court_types_pac),
         judge_id = factor(judge_id),
         year = factor(year, ordered = TRUE))
```


```{r}
# Log transform covariates
dockets_data <- dockets_data %>% 
  mutate(log_age = log(age),
         log_prior_dockets = log(number_prior_dockets + 0.5),
         log_M = log(M + 0.5),
         log_M1 = log(M1 + 0.5),
         log_M2 = log(M2 + 0.5),
         log_M3 = log(M3 + 0.5),
         log_F = log(`F` + 0.5),
         log_F1 = log(F1 + 0.5),
         log_F2 = log(F2 + 0.5),
         log_F3 = log(F3 + 0.5),
         log_S = log(S + 0.5),
         log_S1 = log(S1 + 0.5),
         log_S2 = log(S2 + 0.5),
         log_S3 = log(S3 + 0.5),
         log_IC = log(IC + 0.5),
         log_H1 = log(H1 + 0.5),
         log_H2 = log(H2 + 0.5)
         )
```

```{r}
# create binary response variable
dockets_data <- dockets_data %>% 
  mutate(confinement = if_else(total_confinement_days > 0, 1, 0),
         confinement = factor(confinement))
```


Build a random sample.
```{r}
set.seed(30318)
dockets_sample <- sample_n(dockets_data, size = 1000)
```

```{r}
glimpse(dockets_sample)
```



# 1. Random Effects Model: No transformation on response variables

## 1.1 Dependent variable: Total confinement days

Model specification
*Level 1* $$\text{y}_{ij} = \beta_{0j} + \beta_{1}x_{1} + \dots + \beta_{47}x_{47} + e_{ij}$$


*Level 2* $$\beta_{0j} = \beta_{0} + u_{0j}$$

where $j  = \text{judge_id}$ and $x$s are the covariates.


```{r}
judge_total_intercept_model <- lmer(total_confinement_days ~ 1 + gender + race + age +
                               number_prior_dockets + M + M1 +
                               M2 + M3 + `F` +
                               F1 + F2 + F3 +
                               S + S1 + S2 +
                               S3 + IC + H1 + H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month +
                               (1 | judge_id), dockets_data)
```

```{r}
summary(judge_total_intercept_model)
```

```{r}
ranef(judge_total_intercept_model)$judge_id
```

```{r}
judge_total_ranef <- tibble(ranef(judge_total_intercept_model)$judge_id) %>%
  rename(ranef_intercept = `(Intercept)`) %>% 
  mutate(judge_id = row_number())
```

```{r}
judge_total_ranef
```

```{r}
judge_total_ranef %>% 
  ggplot(aes(x = ranef_intercept)) +
  geom_histogram(binwidth = 250, fill = "#88398A") +
  theme_bw() +
  labs(title = "Judge random effects on intercept",
       subtitle = "Outcome variable: Docket total days of confinement",
       x = "judge random effect") +
  theme(
    panel.grid.minor = element_blank()
  )
```

```{r}
summary(judge_total_ranef$ranef_intercept)
```

Order judges on harshness from high to low.
```{r}
judge_total_ranef %>% 
  arrange(desc(ranef_intercept))
```


```{r}
dockets %>% 
  filter(judge_id == 20) %>% 
  select(disposing_authority__last_name, disposing_authority__first_name,
         disposing_authority__middle_name)
```

### Diagnostics: Total confinement days

fitted vs. residual plot
```{r}
plot(judge_total_intercept_model, type = c("p", "smooth"))
```
Residual are not scattered evenly around zero along the the ranges of fitted values.

scale-location plots
```{r}
plot(judge_total_intercept_model, sqrt(abs(resid(.))) ~ fitted(.), type = c("p", "smooth"))
```
Assumption of homoskedascity is violated.

QQ plot
```{r}
lattice::qqmath(judge_total_intercept_model, id = 0.05)
```
Residuals are not normally distributed.

```{r}
qqnorm(residuals(judge_log_total_intercept_model)); qqline(residuals(judge_log_total_intercept_model))
```


## 1.2 Dependent variable: Max confinement days

```{r}
judge_max_intercept_model <- lmer(max_confinement_days ~ 1 + gender + race + age +
                               number_prior_dockets + M + M1 +
                               M2 + M3 + `F` +
                               F1 + F2 + F3 +
                               S + S1 + S2 +
                               S3 + IC + H1 + H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month +
                               (1 | judge_id), dockets_data)
```


```{r}
judge_max_ranef <- tibble(ranef(judge_max_intercept_model)$judge_id) %>%
  rename(ranef_intercept = `(Intercept)`) %>% 
  mutate(judge_id = row_number())

```

```{r}
judge_max_ranef %>% 
  ggplot(aes(x = ranef_intercept)) +
  geom_histogram(binwidth = 250, fill = "#88398A") +
  theme_bw() +
  labs(title = "Judge random effects on intercept",
       subtitle = "Outcome variable: Docket max days of confinement",
       x = "judge random effect") +
  theme(
    panel.grid.minor = element_blank()
  )
```
Order judges on harshness from high to low.
```{r}
judge_max_ranef %>% 
  arrange(desc(ranef_intercept))
```

### Diagnostics: Max confinement days

fitted vs. residual plot
```{r}
plot(judge_max_intercept_model, type = c("p", "smooth"))
```

scale-location plots
```{r}
plot(judge_max_intercept_model, sqrt(abs(resid(.))) ~ fitted(.), type = c("p", "smooth"))
```

QQ plot
```{r}
lattice::qqmath(judge_max_intercept_model, id = 0.05)
```

# 2. Random Effects Model: Log transformation on response variables

## 2.1 Dependent variable: Log total confinement days

```{r}
judge_log_total_intercept_model <- lmer(log(total_confinement_days + 0.5) ~ 1 + gender + race + age +
                               number_prior_dockets + M + M1 +
                               M2 + M3 + `F` +
                               F1 + F2 + F3 +
                               S + S1 + S2 +
                               S3 + IC + H1 + H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month +
                               (1 | judge_id), dockets_data)
```

```{r}
summary(judge_log_total_intercept_model)
```
```{r}
ranef(judge_log_total_intercept_model)$judge_id
```

```{r}
judge_log_total_ranef <- tibble(ranef(judge_log_total_intercept_model)$judge_id) %>%
  rename(ranef_intercept = `(Intercept)`) %>% 
  mutate(judge_id = row_number())
```

```{r}
judge_log_total_ranef
```

```{r}
judge_log_total_ranef %>% 
  ggplot(aes(x = ranef_intercept)) +
  geom_histogram(fill = "#88398A") +
  theme_bw() +
  labs(title = "Judge random effects on intercept",
       subtitle = "Outcome variable: Log total days of confinement",
       x = "judge random effect") +
  theme(
    panel.grid.minor = element_blank()
  )
```

### Diagnostics: Total confinement days

```{r}
plot(judge_log_total_intercept_model, type = c("p", "smooth"))
```

```{r}
plot(judge_log_total_intercept_model, sqrt(abs(resid(.))) ~ fitted(.), type = c("p", "smooth"))
```

```{r}
lattice::qqmath(judge_log_total_intercept_model, id = 0.05)
```

```{r}
qqnorm(residuals(judge_log_total_intercept_model)); qqline(residuals(judge_log_total_intercept_model))
```

## 2.2 Dependent variable: Log total confinement days, log transformations on some explanatory variables

```{r}
judge_log_total_m2 <- lmer(log(total_confinement_days + 0.5) ~ 1 + gender + race + age +
                               log_prior_dockets + log_M + log_M1 +
                               log_M2 + log_M3 + log_F +
                               log_F1 + log_F2 + log_F3 +
                               log_S + log_S1 + log_S2 +
                               log_S3 + log_IC + log_H1 + log_H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month +
                               (1 | judge_id), dockets_data)
```

```{r}
summary(judge_log_total_m2)
```


```{r}
judge_log_total_ranef_m2 <- tidy_judge_ranef(judge_log_total_m2)
head(judge_log_total_ranef_m2)
```


```{r}
judge_log_total_ranef_m2 %>% 
  ggplot(aes(x = ranef_intercept)) +
  geom_histogram(fill = "#88398A") +
  theme_bw() +
  labs(title = "Judge random effects on intercept",
       subtitle = "Outcome variable: Log total days of confinement",
       x = "judge random effect") +
  theme(
    panel.grid.minor = element_blank()
  )
```

### Diagnostics

```{r}
plot(judge_log_total_m2, type = c("p", "smooth"))
```


```{r}
plot(judge_log_total_m2, sqrt(abs(resid(.))) ~ fitted(.), type = c("p", "smooth"),
     ylab=expression(sqrt(abs(resid))),
     xlab="Fitted values")
```


```{r}
lattice::qqmath(judge_log_total_m2, id = 0.05)
```

```{r}
qqnorm(residuals(judge_log_total_m2), col = "darkgrey"); qqline(residuals(judge_log_total_m2))
```

```{r}
judge_log_total_ranef_m2 %>% 
  ggplot(aes(sample = ranef_intercept)) +
  geom_qq(color = "darkgrey") +
  geom_qq_line() +
  theme_bw() +
  labs(title = "Normal Q-Q Plot: Judge random effects on intercept",
       subtitle = "Outcome variable: Log total days of confinement",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles"
       ) +
  theme(
    panel.grid = element_blank()
  )
```
### Testing significance of random effects

```{r}
judge_log_total_m2_1 <- lmer(log(total_confinement_days + 0.5) ~ 1 + gender + race + age +
                               log_prior_dockets + log_M + log_M1 +
                               log_M2 + log_M3 + log_F +
                               log_F1 + log_F2 + log_F3 +
                               log_S + log_S1 + log_S2 +
                               log_S3 + log_IC + log_H1 + log_H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month +
                               (1 | judge_id), dockets_data, REML = FALSE)

judge_log_total_m2_0 <- lm(log(total_confinement_days + 0.5) ~ 1 + gender + race + age +
                               log_prior_dockets + log_M + log_M1 +
                               log_M2 + log_M3 + log_F +
                               log_F1 + log_F2 + log_F3 +
                               log_S + log_S1 + log_S2 +
                               log_S3 + log_IC + log_H1 + log_H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month, dockets_data)

anova(judge_log_total_m2_1, judge_log_total_m2_0)

```



## 2.3 Dependent variable: Log max confinement days, log transformations on some explanatory variables


```{r}
judge_log_max_m2 <- lmer(log(max_confinement_days + 0.5) ~ 1 + gender + race + age +
                               log_prior_dockets + log_M + log_M1 +
                               log_M2 + log_M3 + log_F +
                               log_F1 + log_F2 + log_F3 +
                               log_S + log_S1 + log_S2 +
                               log_S3 + log_IC + log_H1 + log_H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month +
                               (1 | judge_id), dockets_data)
```

```{r}
summary(judge_log_max_m2)
```

```{r}
judge_log_max_ranef_m2 <- tidy_judge_ranef(judge_log_max_m2)
head(judge_log_max_ranef_m2)
```

```{r}
judge_log_max_ranef_m2 %>% 
  ggplot(aes(x = ranef_intercept)) +
  geom_histogram(fill = "#88398A") +
  theme_bw() +
  labs(title = "Judge random effects on intercept",
       subtitle = "Outcome variable: Log max days of confinement",
       x = "judge random effect") +
  theme(
    panel.grid.minor = element_blank()
  )
```


### Diagnostics
```{r}
plot(judge_log_max_m2, type = c("p", "smooth"))
```

```{r}
plot(judge_log_max_m2, sqrt(abs(resid(.))) ~ fitted(.), type = c("p", "smooth"),
     ylab=expression(sqrt(abs(resid))),
     xlab="Fitted values")
```
```{r}
qqnorm(residuals(judge_log_max_m2), col = "darkgrey"); qqline(residuals(judge_log_max_m2))
```
```{r}
judge_log_max_ranef_m2 %>% 
  ggplot(aes(sample = ranef_intercept)) +
  geom_qq(color = "darkgrey") +
  geom_qq_line() +
  theme_bw() +
  labs(title = "Normal Q-Q Plot: Judge random effects on intercept",
       subtitle = "Outcome variable: Log max days of confinement",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles"
       ) +
  theme(
    panel.grid = element_blank()
  )
```

### Testing significance of random effects

```{r}
judge_log_max_m2_1 <- lmer(log(max_confinement_days + 0.5) ~ 1 + gender + race + age +
                               log_prior_dockets + log_M + log_M1 +
                               log_M2 + log_M3 + log_F +
                               log_F1 + log_F2 + log_F3 +
                               log_S + log_S1 + log_S2 +
                               log_S3 + log_IC + log_H1 + log_H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month +
                               (1 | judge_id), dockets_data, REML = FALSE)

judge_log_max_m2_0 <- lm(log(max_confinement_days + 0.5) ~ 1 + gender + race + age +
                               log_prior_dockets + log_M + log_M1 +
                               log_M2 + log_M3 + log_F +
                               log_F1 + log_F2 + log_F3 +
                               log_S + log_S1 + log_S2 +
                               log_S3 + log_IC + log_H1 + log_H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month, dockets_data)

anova(judge_log_max_m2_1, judge_log_max_m2_0)

```


# 3. Random Effects Model: inverse gaussian GLM, link = log

## Dependent variable: Total confinement days

```{r}
# This chuck generates error below:
# Error in (function (fr, X, reTrms, family, nAGQ = 1L, verbose = 0L, maxit = 100L, : (maxstephalfit) PIRLS step-halvings failed to reduce deviance in pwrssUpdate

judge_iG_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                               number_prior_dockets + M + M1 +
                               M2 + M3 + `F` +
                               F1 + F2 + F3 +
                               S + S1 + S2 +
                               S3 + IC + H1 + H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month +
                               (1 | judge_id), data = dockets_data,
                               family = inverse.gaussian(link = "log"))
```

```{r}
# This chunk kept on running and was terminated.
# log_S1, log_S2, log_S3, and log_IC are excluded because they vary very little, their 
# inclusion would result in glmer warning : “Downdated VtV is not positive definite”
judge_iG_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + log_age +
                                   log_prior_dockets + log_M + log_M1 + log_M2 + log_M3 + log_F +
                                   log_F1 + log_F2 + log_F3 + log_S + 
                                   log_H1 + log_H2 + court_types_cp + court_types_mc +
                                   court_types_pac + year + month + 
                                   (1 | judge_id), data = dockets_data,
                                 family = inverse.gaussian(link = "log"))
```
```{r}
# This chunk kept on running and was terminated.
# log_S1, log_S2, log_S3, and log_IC are excluded because they vary very little, their 
# inclusion would result in glmer warning : “Downdated VtV is not positive definite”
judge_iG_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                                   log_prior_dockets + log_M + log_M1 + log_M2 + log_M3 + log_F +
                                   log_F1 + log_F2 + log_F3 + log_S + 
                                   log_H1 + log_H2 + court_types_cp + court_types_mc +
                                   court_types_pac + year + month + 
                                   (1 | judge_id), data = dockets_data,
                                 family = inverse.gaussian(link = "log"))
```

```{r}
# This chunk generates error below:
# Error in (function (fr, X, reTrms, family, nAGQ = 1L, verbose = 0L, maxit = 100L, : Downdated VtV is not positive definite
judge_iG_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                                   log_prior_dockets +  log_M1 + log_M2 + log_M3 + 
                                   log_F1 + log_F2 + log_F3 + log_S + 
                                   log_H1 + log_H2 + court_types_cp + court_types_mc +
                                   court_types_pac + year + month + 
                                   (1 | judge_id), data = dockets_data,
                                 family = inverse.gaussian(link = "log"))
```
### run on docket samples

```{r}
# This chunk generates error below:
# Error in (function (fr, X, reTrms, family, nAGQ = 1L, verbose = 0L, maxit = 100L, : pwrssUpdate did not converge in (maxit) iterations

judge_iG_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                                   log_prior_dockets + log_M + log_M1 + log_M2 + log_M3 + log_F +
                                   log_F1 + log_F2 + log_F3 + log_S + log_IC +
                                   log_H1 + log_H2 + court_types_cp + court_types_mc +
                                   court_types_pac + year + month + 
                                   (1 | judge_id), data = dockets_sample,
                                 family = inverse.gaussian(link = "log"))
```

```{r}
# This chunk generates error below:
#Error in (function (fr, X, reTrms, family, nAGQ = 1L, verbose = 0L, maxit = 100L, : pwrssUpdate did not converge in (maxit) iterations
judge_iG_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                                   log_prior_dockets + log_M1 + log_M2 + log_M3 + 
                                   log_F1 + log_F2 + log_F3 + log_S + log_IC +
                                   log_H1 + log_H2 + court_types_cp + court_types_mc +
                                   court_types_pac + year + month + 
                                   (1 | judge_id), data = dockets_sample,
                                 family = inverse.gaussian(link = "log"))
```

# 4. Random Effects Model: gamma GLM, link = log

```{r}
# This chunk kept on running and was terminated.
# Warning message:Some predictor variables are on very different scales: consider rescaling.
judge_gamma_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                               number_prior_dockets + M + M1 +
                               M2 + M3 + `F` +
                               F1 + F2 + F3 +
                               S + S1 + S2 +
                               S3 + IC + H1 + H2 +
                               court_types_cp + court_types_mc + court_types_pac +
                               year + month +
                               (1 | judge_id), data = dockets_data,
                               family = Gamma(link = "log"))
```
```{r}
# This chunk kept on running and was terminated.
judge_gamma_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                                   log_prior_dockets + log_M + log_M1 + log_M2 + log_M3 + log_F +
                                   log_F1 + log_F2 + log_F3 + log_S + 
                                   log_H1 + log_H2 + court_types_cp + court_types_mc +
                                   court_types_pac + year + month + 
                                   (1 | judge_id), data = dockets_data,
                                   family = Gamma(link = "log"))
```

```{r}
# This chunk was terminated
judge_gamma_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                                   log_prior_dockets + log_M + log_M1 + log_M2 + log_M3 + log_F +
                                   log_F1 + log_F2 + log_F3 + log_S + 
                                   log_H1 + log_H2 + year + month + 
                                   (1 | judge_id), data = dockets_data,
                                   family = Gamma(link = "log"))
```
```{r}
# This chunk was terminated
judge_gamma_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + 
                                   log_prior_dockets + log_M + log_M1 + log_M2 + log_M3 + log_F +
                                   log_F1 + log_F2 + log_F3 + log_S + 
                                   log_H1 + log_H2 + year + month + 
                                   (1 | judge_id), data = dockets_data,
                                   family = Gamma(link = "log"))
```

```{r}
#This chunk was terminated
judge_gamma_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + 
                                   log_prior_dockets + year + month + 
                                   (1 | judge_id), data = dockets_data,
                                   family = Gamma(link = "log"))
```

### run on docket samples

```{r}
# This chunk generates error below:
# fixed-effect model matrix is rank deficient so dropping 1 column / coefficient
# failure to converge in 10000 evaluationsconvergence code 4 from Nelder_Mead: failure to converge 
# in 10000 evaluationsModel failed to converge with max|grad| = 1.06665 (tol = 0.002, component 1)
judge_gamma_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                                   log_prior_dockets + log_M + log_M1 + log_M2 + log_M3 + log_F +
                                   log_F1 + log_F2 + log_F3 + log_S + log_IC +
                                   log_H1 + log_H2 + court_types_cp + court_types_mc +
                                   court_types_pac + year + month + 
                                   (1 | judge_id), data = dockets_sample,
                                 family = Gamma(link = "log"))
```

```{r}
# This chunk generates error below:
# fixed-effect model matrix is rank deficient so dropping 1 column / coefficient
# failure to converge in 10000 evaluationsconvergence code 4 from Nelder_Mead: failure to converge 
# in 10000 evaluationsModel failed to converge with max|grad| = 0.616358 (tol = 0.002, component 1)
judge_gamma_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                                   log_prior_dockets + log_M + log_M1 + log_M2 + log_M3 + log_F +
                                   log_F1 + log_F2 + log_F3 + log_S + 
                                   log_H1 + log_H2 + court_types_cp + court_types_mc +
                                   court_types_pac + year + month + 
                                   (1 | judge_id), data = dockets_sample,
                                 family = Gamma(link = "log"))
```

```{r}
# This chunk generates error below:
# fixed-effect model matrix is rank deficient so dropping 1 column / coefficient
# failure to converge in 10000 evaluationsconvergence code 4 from Nelder_Mead: failure to converge 
# in 10000 evaluationsModel failed to converge with max|grad| = 0.29911 (tol = 0.002, component 1)
judge_gamma_total_link_log <- glmer((total_confinement_days + 0.5) ~ 1 + gender + race + age +
                                   log_prior_dockets + log_M1 + log_M2 + log_M3 + 
                                   log_F1 + log_F2 + log_F3 +  
                                   log_H1 + log_H2 + 
                                   year +
                                   (1 | judge_id), data = dockets_sample,
                                 family = Gamma(link = "log"))
```

# 5. Random Effects model: binomial GLM, link = logit

## run on docket samples
```{r}
judge_binom_link_log <- glmer(confinement ~ 1 + gender + race + age +
                                   log_prior_dockets + log_M + log_M1 + log_M2 + log_M3 + log_F +
                                   log_F1 + log_F2 + log_F3 + log_S + log_IC +
                                   log_H1 + log_H2 + court_types_cp + court_types_mc +
                                   court_types_pac + year + month + 
                                   (1 | judge_id), data = dockets_sample,
                                 family = binomial(link = "logit"))
```

```{r}
judge_binom_link_log <- glmer(confinement ~ 1 + gender + race + age +
                                number_prior_dockets + M + M1 + M2 + M3 + `F` + F1 +
                                F2 + F3 + S + IC + H1 + H2 + court_types_cp + court_types_mc +
                                court_types_pac + year + month + 
                                (1 | judge_id), data = dockets_sample,
                                 family = binomial(link = "logit"))
```

