---
title: "R Notebook"
output: html_notebook
---

Exploring the `grades` column.

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(rprojroot))
```
```{r}
source(file.path(rprojroot::find_root(rprojroot::is_rstudio_project),
                 "analyses", "team2", "preprocess", "ec_functions.R"))
```


In this notebook I'm only going to examine data that are present in the `offenses_dispositions` data file. 

```{r}
dispositions <- read_csv(file.path(find_root(is_rstudio_project), "data",
                                   "offenses_dispositions_v3_periods.csv"),
                         col_types = cols(
  X1 = col_double(),
  docket_id = col_double(),
  description = col_character(),
  statute_description = col_character(),
  statute_name = col_character(),
  sequence_number = col_double(),
  grade = col_character(),
  disposition = col_character(),
  disposing_authority__first_name = col_character(),
  disposing_authority__middle_name = col_character(),
  disposing_authority__last_name = col_character(),
  disposing_authority__title = col_character(),
  disposing_authority__document_name = col_character(),
  disposition_method = col_character(),
  min_period = col_character(),
  max_period = col_character(),
  period = col_character(),
  credit = col_double(),
  sentence_type = col_factor(levels = c("No Further Penalty",
                                        "Merged",
                                        "Probation",
                                        "IPP",
                                        "Confinement")),
  min_period_parsed = col_character(),
  max_period_parsed = col_character(),
  min_period_days = col_double(),
  max_period_days = col_double()
))
```

Let's look at grades and sentence types.

```{r}
dispositions %>%
  filter(!is.na(sentence_type), !is.na(grade)) %>%
  ggplot(aes(sentence_type)) +
  geom_bar() +
  coord_flip() +
  facet_wrap(~ grade, scales = "free_x") +
  theme_minimal()
```

I don't know what's up with the `grade` values of H1, H2, IC, and I also don't know what's going on with S, M, and F without a number.

```{r}
dispositions %>%
  filter(grade %in% c("H1", "H2")) %>%
  count(grade, statute_name) %>%
  arrange(grade, desc(n)) %>%
  group_by(grade) %>%
  slice(1:5)
```

H stands for "homicide", 18 § 2502 §§ A is murder of the first degree and 18 § 2502 §§ B is murder of the second degree (which could have been coded as a F1, but was not). 

```{r}
dispositions %>%
  filter(statute_name == "18 § 2502 §§ B") %>%
  count(grade)
```

What about "IC"?

```{r}
dispositions %>%
  filter(grade == "IC") %>%
  count(statute_name) %>%
  arrange(desc(n)) %>%
  head(10)
```

These are cases of contempt for violating a protection order.

# Cleaning up `NA`s.

How often is `grade` set to `NA` but a `statute_name` is available, anyway?

```{r}
# Clean up this error
dispositions <- dispositions %>%
  mutate(statute_name = str_replace_all(statute_name, "ยง", "§"))

dispositions %>%
  filter(is.na(grade),
         !is.na(statute_name)) %>%
  nrow()
```

Whoa, that's a lot. 

For each statute name, find the most common grade, and the percentage of instances of that statute for which that grade is selected.

```{r}
most_popular_grade <- dispositions %>%
  filter(!is.na(grade), !is.na(statute_name)) %>%
  count(grade, statute_name, name = "num_dispositions") %>%
  arrange(statute_name, desc(num_dispositions)) %>%
  group_by(statute_name) %>%
  mutate(statute_pct = 100 * num_dispositions / sum(num_dispositions))

most_popular_grade %>%
  slice(1) %>%
  ungroup()  %>%
  arrange(statute_pct) %>%
  mutate(cum_density = row_number(statute_pct) / nrow(.)) %>%
  ggplot(aes(statute_pct, cum_density)) +
  geom_line() +
  theme_minimal()
```

This figure tells us that, for 75% of the `statute_name` values, a single `grade` accounts for 100% of the dispositions. We can use these cases to fill in missing `grade` values.

```{r}
top_most_popular_grade <- most_popular_grade %>%
  filter(statute_pct > 100 - 1e-8) %>%
  select(statute_name, most_common_grade = grade)
  
dispositions <- left_join(dispositions, top_most_popular_grade, 
                          by = "statute_name") %>%
  mutate(new_grade = ifelse(is.na(grade), most_common_grade, grade))

dispositions %>%
  filter(is.na(new_grade),
         !is.na(statute_name)) %>%
  nrow()
```

This step recovers tens of thousands of missing `grade` values, so it's probably worth making this process into a function. 

...done.

Now make sure that the function version returns the same thing.

```{r}
all.equal(dispositions$new_grade,
          backfill_disposions_grades(dispositions$grade, 
                                     dispositions$statute_name))
```

Now let's see what's going on when grade "S", "M", and "F" don't have an associated number. Are these just missing data?

Find the top ten statutes for each grade and compare them.

```{r}
most_popular_statute <- most_popular_grade %>%
  group_by(grade) %>%
  mutate(grade_pct = 100 * num_dispositions / sum(num_dispositions),
         rank = row_number(100 - grade_pct)) %>%
  filter(rank <= 10) %>%
  mutate(statute_name = paste0(statute_name, 
                               " (", format(num_dispositions, big.mark = ","), ")")) %>%
  select(statute_name, rank, grade) %>%
  arrange(grade, rank) %>%
  pivot_wider(names_from = "grade", values_from = "statute_name")

select(most_popular_statute, starts_with("F"))
```
```{r}
select(most_popular_statute, starts_with("M"))
```
```{r}
select(most_popular_statute, starts_with("S"))
```

To my (non-expert) eye, these are all different enough, but have enough overlap, to conclude that `statute_name` and `grade` provide different information, and that grades of "M" and "F", without a number, are meaningful and not errors.
