---
title: "Preprocess Cleaned Data"
author: "Shanti Agung"
date: "4/22/2021"
output: html_document
---

This notebook is to explore ways to preprocess the cleaned dataset in order to create variables needed for modeling.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
# load dataset
dispositions <- read_csv(file.path(rprojroot::find_root(rprojroot::is_rstudio_project),
                                   "data", "offenses_dispositions_v3_grades.csv"),
                         col_types = cols(
  #X1 = col_double(),
  docket_id = col_double(),
  description = col_character(),
  statute_description = col_character(),
  statute_name = col_character(),
  sequence_number = col_double(),
  grade = col_character(),
  disposition = col_character(),
  disposing_authority__first_name = col_character(),
  disposing_authority__middle_name = col_character(),
  disposing_authority__last_name = col_character(),
  disposing_authority__title = col_character(),
  disposing_authority__document_name = col_character(),
  disposition_method = col_character(),
  min_period = col_character(),
  max_period = col_character(),
  period = col_character(),
  credit = col_double(),
  sentence_type = col_character(),
  min_period_parsed = col_character(),
  max_period_parsed = col_character(),
  min_period_days = col_character(),
  max_period_days = col_double(),
  grade_backfilled = col_character()
))


```

```{r}
dddcols = cols(
  docket_id = col_double(),
  gender = col_character(),
  race = col_character(),
  date_of_birth = col_date(format = ""),
  arrest_date = col_date(format = ""),
  complaint_date = col_date(format = ""),
  disposition_date = col_date(format = ""),
  filing_date = col_date(format = ""),
  initiation_date = col_date(format = ""),
  status_name = col_character(),
  court_office__court__display_name = col_character(),
  current_processing_status__processing_status = col_character(),
  current_processing_status__status_change_datetime = col_date(format = ""),
  municipality__name = col_character(),
  municipality__county__name = col_character(),
  judicial_districts = col_character(),
  court_office_types = col_character(),
  court_types = col_character(),
  representation_type = col_character()
)
dockets <- readr::read_csv('https://storage.googleapis.com/jat-rladies-2021-datathon/defendant_docket_details.csv',
                       col_types = dddcols)
```

```{r}
ddidscols = cols(
  defendant_id = col_double(),
  docket_id = col_double()
)

dockets_defendants_ids <- readr::read_csv('https://storage.googleapis.com/jat-rladies-2021-datathon/defendant_docket_ids.csv',
                                      col_types = ddidscols)

```

## Unit of analysis

Unit of analysis will be `docket_id`, rather than offense,  because judges seem to make sentencing decision based on overall bundle of offenses within a docket, rather than individual offense. One example of this decision making process is that some offenses are merged.

## Target (dependent) variables

Several possible target variables:
1. Days of confinement sentences (`total_confinement_days` and `max_confinement_days` are defined below)
2. Binary: confinement and non-confinement
3. ...

### Days of confinement 

```{r}
dispositions %>% 
  count(sentence_type)
```

Aside from "Confinement", there are four other values of `sentence_type`. By defining days of confinement as the dependent variable:
* `days_of_confinement` is the value of `max_period_days` if `sentence_type` is "Confinement".
* `days_confinement` is 0 for all other values of `sentence_type`.

How many "Confinement" are there in a docket?
```{r}
dispositions %>% 
  mutate(is_confinement = if_else(sentence_type == "Confinement", TRUE, FALSE)) %>% 
  group_by(docket_id) %>% 
  summarise(n_confinement = sum(is_confinement)) %>% 
  ungroup() %>% 
  count(n_confinement)
```
The majority of dockets do not result in confinement. Some dockets list more than one confinement as `sentence_type` of its offenses. Let's take a look at some of these dockets.

```{r}
dispositions %>% 
  mutate(is_confinement = if_else(sentence_type == "Confinement", TRUE, FALSE)) %>% 
  group_by(docket_id) %>% 
  summarise(n_confinement = sum(is_confinement)) %>% 
  ungroup() %>% 
  filter(n_confinement == 4)
```
`docket_id` number 3 has two offenses that results in confinement. Let's look closer. 
```{r}
dispositions %>% 
  filter(docket_id == 3) 
```
`docket_id` = 3 has 3 offenses, two of them results in confinement. One for 700 days and the other for 1826 days.

Let's look at another `docket_id`. `docket_id` = 104 has 4 offenses that results in confinement. 
```{r}
dispositions %>% 
  filter(docket_id == 104)
```
This defendant offenses are: conspiracy, carry firearms, criminal attempt, firearms to be carried without license. For each of these offenses, the defendant received different length of confinement. 

_approach_ : for a `docket_id` that has multiple confinement decisions, the `days_of_confinement` is the sum of `max_period_days` for offenses, in that docket, that results in confinement. With this approach, we treat as if all sentences are executed consecutively, rather than concurrently or a combination of them. `total_confinement_days` seems to better variable name.

_approach_: for a `docket_id` that has multiple confinement decisions, the `max_confinement_days` is the max of `max_period_days` for offenses, in that docket, that results in confinement. With this approach, we treat as if all sentences are executed concurrently.

_assumption_: for now, "life" sentence is treated as confinement of 90 years, regardless whether the sentence is life with or without parole. 1 year = 365.2425 (as calculated in ec_sentence_period.Rmd).

Other exclusion, for now:
* We use dockets that the value of `credit` is missing and the value of `max_period_parsed` does not include "time_served" (otherwise, these observations have -Inf as their `max_period_days`)
* We use dockets that the value of `max_period_parsed` is not "other" (otherwise, these observations have -Inf as their `max_period_days`)
* We use dockets that the value of `max_period_parsed` is not "death penalty imposed" (otherwise, these observations have -Inf as their `max_period_days`). There are only two docket_ids receiving this sentence.

```{r}
# create total_confinement_days and max_confinement_days
dockets_confinement <- dispositions %>% 
  filter(!is.na(sentence_type)) %>%
  mutate(has_time_served = str_detect(max_period_parsed, "time_served"),
         missing_credit = is.na(credit),
         has_max_period_parsed = !is.na(max_period_parsed)) %>% 
  filter(!(has_max_period_parsed & has_time_served & missing_credit)) %>% 
  filter((!has_max_period_parsed) | (max_period_parsed != "other")) %>% 
  filter((!has_max_period_parsed) | (max_period_parsed != "death penalty imposed")) %>% 
  mutate(max_period_days = if_else(str_detect(max_period_parsed, "life"), (90*365.2425),
                                   max_period_days)) %>% 
  group_by(docket_id) %>%
  mutate(confinement_days = if_else(sentence_type == "Confinement", max_period_days, 0)) %>% 
  mutate(total_confinement_days = sum(confinement_days, na.rm = TRUE),
         max_confinement_days = max(confinement_days, na.rm = TRUE)) %>% 
  ungroup() %>% 
  select(docket_id, total_confinement_days, max_confinement_days) %>% 
  distinct(docket_id, .keep_all = TRUE)


```


```{r}
dockets_confinement
```

Note that if a docket has all its offenses' `sentence_type` value missing (`NA`), the `docket_id` will not be included in `dockets_confinement`.


## Create covariates

### Number of a grade offenses

Are there offenses with `grade_backfilled` missing? Yes, there are.
```{r}
dispositions %>% 
  count(grade_backfilled)
```

Exclude offenses that have missing values in `grade_backfilled`
```{r}
dispositions %>% 
  select(docket_id, grade_backfilled) %>% 
  filter(!is.na(grade_backfilled)) %>% 
  count(docket_id, grade_backfilled)
```



```{r}
dockets_grades <- dispositions %>% 
  select(docket_id, grade_backfilled) %>% 
  filter(!is.na(grade_backfilled)) %>% 
  count(docket_id, grade_backfilled) %>% 
  pivot_wider(names_from = grade_backfilled, values_from = n) %>% 
  mutate(across(2:16, ~replace_na(., 0)))

```

```{r}
dockets_grades
```


### Judge ID

How many judges are there in the data set?

```{r}
dispositions %>% 
  distinct(disposing_authority__first_name, disposing_authority__middle_name, 
           disposing_authority__last_name) %>% 
  filter(!is.na(disposing_authority__last_name)) %>% 
  select(disposing_authority__last_name, disposing_authority__first_name,
         disposing_authority__middle_name) %>% 
  arrange(disposing_authority__last_name, disposing_authority__first_name) %>% 
  count()
```

Note that there are these three entries in `disposing_authority__last_name`.
```{r}
dispositions %>% 
  distinct(disposing_authority__first_name, disposing_authority__middle_name, 
           disposing_authority__last_name) %>% 
  select(disposing_authority__last_name, disposing_authority__first_name,
         disposing_authority__middle_name) %>% 
  arrange(disposing_authority__last_name, disposing_authority__first_name) %>% 
  filter(str_detect(disposing_authority__last_name, pattern = "Pennsylvania"))
```
```{r}
# create judge_id only
dispositions %>% 
  filter(!is.na(disposing_authority__last_name)) %>% 
  arrange(disposing_authority__last_name, disposing_authority__first_name) %>% 
  mutate(judge_id = group_indices(., disposing_authority__last_name,
                                  disposing_authority__middle_name,
                                  disposing_authority__first_name)) %>% 
  select(judge_id, disposing_authority__last_name, disposing_authority__first_name, 
           disposing_authority__middle_name) %>% 
  distinct(judge_id, .keep_all = TRUE) %>% 
  arrange(judge_id)
  
```
Same number of judges: 191.


```{r}
# create docket - judge_id
# note that docket_judge_id contains only dockets that have non-missing judge's last name
dockets_judges <- dispositions %>% 
  filter(!is.na(disposing_authority__last_name)) %>% 
  arrange(disposing_authority__last_name, disposing_authority__first_name) %>% 
  mutate(judge_id = group_indices(., disposing_authority__last_name,
                                  disposing_authority__middle_name,
                                  disposing_authority__first_name)) %>% 
  select(docket_id, judge_id, disposing_authority__last_name, disposing_authority__first_name, 
           disposing_authority__middle_name) %>% 
  distinct(docket_id, .keep_all = TRUE) %>% 
  arrange(docket_id)
```

```{r}
dockets_judges
```

How many dockets do not have judges?
```{r}
dockets %>% 
  anti_join(dockets_judges) %>% 
  count()
```


### Indicators: court_types

Take a look at different values of `court_types`
```{r}
dockets %>% 
  count(court_types)
```

```{r}
dockets %>% 
  mutate(court_types_cp = if_else(str_detect(court_types, "CP"), 1, 0)) %>% 
  mutate(court_types_mc = if_else(str_detect(court_types, "MC"), 1, 0)) %>% 
  mutate(court_types_pac = if_else(str_detect(court_types, "PAC"), 1, 0)) 
```

### Indicators: court_office_types

Take a look at different values of `court_office_types`
```{r}
dockets %>% 
  count(court_office_types)
```

```{r}
dockets %>% 
  mutate(court_office_types_commonwealth = if_else(str_detect(court_office_types, "Commonwealth"), 1, 0)) %>% 
  mutate(court_office_types_criminal = if_else(str_detect(court_office_types, "Criminal"), 1, 0)) %>% 
  mutate(court_office_types_municipal = if_else(str_detect(court_office_types, "Municipal"), 1, 0)) %>% 
  mutate(court_office_types_supreme = if_else(str_detect(court_office_types, "Supreme"), 1, 0)) %>%
  mutate(court_office_types_suprerior = if_else(str_detect(court_office_types, "Superior"), 1, 0)) 
```


### Indicators: representation_type

```{r}
dockets %>% 
  count(representation_type)
```

### Defendant Age

```{r}
dockets %>% 
  mutate(age = lubridate::year(filing_date) - lubridate::year(date_of_birth))
```

### Case duration

```{r}
dockets %>% 
  mutate(duration = difftime(disposition_date, filing_date, units = "days")) %>% 
  count(duration)
```

```{r}
dockets %>% 
  mutate(docket_duration_days = as.numeric(difftime(disposition_date, filing_date, units = "days"))) %>% 
  count(docket_duration_days)
```
Some docket_ids have duration less than 0, and many has duration 0 days. `disposition_date` may not be the proper variable to create `duration`. It should be `duration` = `termination_date` - `filing_date`. But we do not have `termination_date` variable. 

### Number of prior court appearance

```{r}
dockets_prior <- dockets %>%
  left_join(dockets_defendants_ids) %>% 
  select(defendant_id, docket_id, filing_date) %>% 
  group_by(defendant_id) %>% 
  arrange(defendant_id, filing_date) %>% 
  mutate(tmp1 = row_number(),
         number_prior_dockets = tmp1 - 1) %>% 
  ungroup() %>% 
  select(docket_id, number_prior_dockets)
```
```{r}
dockets_prior
```


### Notes on covariates chosen
Has no missing values:
* `court_types` 
* `judicial_districts`
* `court_office_types`
* `representation_type`

Information may be captured by judicial districts, and has missing values:
* `municipality_name`
* `municipality__county__name`


## Create docket dataset for analysis

```{r}
# check initial number of dockets
dim(dockets)
```


```{r}
dockets_data <- dockets %>% 
  left_join(dockets_grades) %>% 
  left_join(dockets_judges) %>% 
  left_join(dockets_prior) %>% 
  left_join(dockets_confinement) %>% 
  mutate(age = lubridate::year(filing_date) - lubridate::year(date_of_birth)) %>% 
  mutate(court_types_cp = if_else(str_detect(court_types, "CP"), 1, 0)) %>% 
  mutate(court_types_mc = if_else(str_detect(court_types, "MC"), 1, 0)) %>% 
  mutate(court_types_pac = if_else(str_detect(court_types, "PAC"), 1, 0)) %>% 
  mutate(court_office_types_commonwealth = if_else(str_detect(court_office_types, "Commonwealth"), 1, 0)) %>% 
  mutate(court_office_types_criminal = if_else(str_detect(court_office_types, "Criminal"), 1, 0)) %>% 
  mutate(court_office_types_municipal = if_else(str_detect(court_office_types, "Municipal"), 1, 0)) %>% 
  mutate(court_office_types_supreme = if_else(str_detect(court_office_types, "Supreme"), 1, 0)) %>%
  mutate(court_office_types_suprerior = if_else(str_detect(court_office_types, "Superior"), 1, 0)) 
```

```{r}
# check dimension
dim(dockets_data)
```

```{r}
# Make sure the target file (docket details data) does not exist
target_file <- file.path(rprojroot::find_root(rprojroot::is_rstudio_project),
                         "data", "docket_details.csv")
stopifnot(!file.exists(target_file))
```

```{r}
# Write it to a new file
readr::write_csv(dockets_data, target_file)
```




```


