---
title: "Bail Race"
author: "Katrina Gutierrez"
date: "3/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Purpose of this file: 

- explore bail data issues (multiple initial bail amounts and multiple judges)
- merge bail with defendant and offenses data 
- look at bail trends disaggretated by race and the grade of the offense


## Initial set-up

Loading libraries

```{r }
library("dplyr")
library("ggplot2")
library("lubridate")
library("DataExplorer")

setwd("C:/Users/katri/Documents/GT/R-Ladies Datathon 2021")
```


Loading the full datasets (all years):

- bail: each record is an action related to bail; each docket_id can have multiple actions
- off: offense and disposition data; each record is an offense/disposition; each docket_id can have multiple offenses
- def: defendent docket data; each record is one defendent/docket_id; each docket_id has one defendent

```{r }

bail <- read.csv("bail.csv")
off  <- read.csv("offenses_dispositions.csv")
def  <- read.csv("defendant_docket_details.csv")

```


## Cleaning Bail Data

An initial look at the bail data
```{r }
glimpse(bail)
```

Since I want to get initial bail for each docket_id and there are many actions for a given date (month), I want "Set" to be the first action type for each date when I sort the data later.  Therefore, I made the variable action_type_name into a factor and put "Set" as the first level.

Even if "Set" is not in the first month, we will still be able to get whatever bail amount is included in the first month.

```{r }

#releveling action_type_name so that "Set" comes first 
bail$action_type_name <- factor(bail$action_type_name)
bail$action_type_name <- relevel(bail$action_type_name, ref = "Set")

#checking the levels for action_type_name
levels(bail$action_type_name)

```

Initial Cleaning Actions:

- filtered for participant title "Judge"
- created a variable "judge" with the first and last name
- converted action_date to a date type
- sorted by docket_id, then action_date, and then action_type_name

```{r }

bail <- bail %>% 
  filter(participant_name__title=="Judge",!is.na(participant_name__first_name))  %>% 
  mutate(judge=paste(participant_name__last_name,participant_name__first_name,sep=","))%>% 
  mutate(action_date = ymd(action_date))%>% 
  mutate(action_type_name = factor(action_type_name))%>% 
  arrange(docket_id, action_date, action_type_name)

```

The action date column uses the first day of the month, so it is difficult to determine the order of events that occurred within the same month. It is also difficult to determine what the initial bail is if there are multiple values for bail in the first month.

To explore this issue, I first looked at how many bail amounts were associated with a given action date and action type for each docket.  In most cases, there was only one value for bail for a given action date + action type + docket id. However, in some cases there were up to 3 different bail amounts in the same month for a docket, and these also had the same action type. 

```{r }

#Grouped the bail data to see the number unique bail amounts associated with a given docket_id + action_date + action_type_name
test1 <- bail %>% group_by(docket_id, action_date, action_type_name)%>% summarise(
  n_bail_vals = length(unique(total_amount)))

#Usually there is only one value for bail, but sometimes there's up to 3.  Unclear which would come first in those cases.
summary(test1$n_bail_vals)

```

Since I wanted initial bail for each docket, I looked at how severe this issue was for the first action date in particular per docket. 

Most dockets had only one value for bail amount for the "Set" type of actions in the first month. However, there were a few that had 2 different bail amounts in the first month and both with a "Set" action type.

```{r }

#Grouping the result of the previous test by docket_id to see what is happening for the first action date.  The data was arleady sorted by action date previously and with action type "Set" at the top for each date.
test2 <- test1 %>% group_by(docket_id)%>% summarise(
  n_bail_vals = first(n_bail_vals))

#For the first action date, there is usually only one value for bail.  But sometimes there are two.
summary(test2$n_bail_vals)

```

Next, I wanted a better understanding of how frequently there are two different bail values being set in first month. 

It turns out this issue is pretty rare.  Meaning, there are 188 docket_ids (0.054% of cases for all years) where there are two bail values to choose from for initial bail.

```{r }

#filters the previous results for dockets with more than one value for bail in the first month
test3 <- test2 %>% filter(n_bail_vals > 1)

#Creates a list of the docket ids fitting this criteria
mult_inital_bail <- unique(test3$docket_id)

#This issue rare:  188 docket ids for all years
length(mult_inital_bail)

#In other words, 0.054% of cases for all years
length(mult_inital_bail)/length(unique(bail$docket_id))

```

Here, we take a look at some of the dockets that have two bail values in the first month.  It's unclear which value would be first.

*We still need to figure out what to do about these cases*

```{r }

#filters data for docket ids that have more than one bail value in the first month based off the previous code cell
mult_inital_bail_data <- bail %>% filter(docket_id %in% mult_inital_bail )
head(mult_inital_bail_data)

```

Creating a dataset where each record is a docket_id with summary bail info like:

- number of unique judges associated with the case
- minimum/maximum bail ever set during the case
- first and last bail amount set + the dates

```{r }

#grouping and summarizing the data, such that we have one record per docket_id for bail
bailGrouped <- group_by(bail, docket_id) %>% summarise(
  n_judges = length(unique(judge)),
  min_bail = min(total_amount),
  max_bail = max(total_amount),
  date_first_bail = first(action_date),
  first_bail = first(total_amount),
  date_last_bail = last(action_date),
  last_bail = last(total_amount))


```

Now, I'll take a quick look at the distribution of the number of judges per case. While most cases have less than 5 judges, there can be up to 20.

```{r }
summary(bailGrouped$n_judges)

```

It's difficult to understand what each judge has done for each docket because all of the judges that are involved are connected with each action date, type, and corresponding bail amount.

So for example, if we wanted to look at action type "Increase Bail Amount" on day X for docket Y, we would see all of the judges associated with docket Y for that action.

```{r }

bail <- bail %>% arrange(docket_id, action_date, action_type_name, total_amount, registry_entry_code)

#grouping the data such that we can see how many unique judges are associated with a particular action type and amount of bail within one action date per docket 
judgeGrouped <- bail %>% group_by(docket_id, action_date, action_type_name, total_amount) %>% summarise(
  n_judges = length(unique(judge)))

#getting the number of unique judges for each docket 
judgeGrouped2 <- bail %>% group_by(docket_id) %>% summarise(
  n_judges2 = length(unique(judge)))

#merging these two things
joinedJudge <- left_join(judgeGrouped, judgeGrouped2, by = "docket_id")

#creating a variable where we see if the number of unique judges for an action type is the same as the number of unique judges for the whole docket in general
joinedJudge$same <- ifelse(joinedJudge$n_judges == joinedJudge$n_judges2, 1, 0)

#it turns out that for all action types for each date within a docket, it has all the judges associated with the docket in general (because all the values are ones)
summary(joinedJudge$same )

```

If we were to look at the actions on an even more granular level with the registry entry code (for the same docket, date, action type, bail amount), we would find different judges associated with different values of registry entry code. Usually, at this granular level, there is only one judge associated with each registry entry code, but there can be up to 17.

I'm not sure how useful this granular data is though because there are 1508 registry entry codes, and I don't know what exactly they mean.  

```{r }
#there are 1508 different registry entry codes
length(unique(bail$registry_entry_code))

judgeGrouped3 <- bail %>% group_by(docket_id, action_date, action_type_name, total_amount, registry_entry_code) %>% summarise(
  n_judges = length(unique(judge)))

#most of the time only one judge is associated with one registry entry code, but there can be up to 17 judges
summary(judgeGrouped3$n_judges)

```

Since some cases have multiple judges and it's difficult to distinguish their involvement, it's difficult to look at bail trends with judges.


## Defendent Data

For the defendant data, each unique docket_id is associated with one record.  However, there's more docket ids in the defendant data than we have for bail.

```{r }
#number of records in defendant data = number of unique docket ids
nrow(def) == length(unique(def$docket_id))

length(unique(def$docket_id)) #number of docket ids for defendant data
length(unique(bailGrouped$docket_id)) #number of docket ids for bail data

```

An overview of the defendant data

```{r }
glimpse(def)

```

Overview of frequency of cases per race.  

I recognize this is not the prettiest graph, but I have a feeling the team looking at defendant data has more graphs and insights.

Regardless, most cases have either black or white defendants, so I chose to compare black and white in my analysis later.

```{r }
defRace <- group_by(def, race) %>% summarise(
  n_cases = length(docket_id))

#plots a bar chart with the number of cases per offense grade
Cases_Race <- ggplot(defRace, aes(x=race, y=n_cases, fill = race)) + 
  geom_bar(stat = "identity") + 
  labs(title="Number of Cases by Race",  
       y="Num Cases", x="Race") +
  theme(axis.ticks.x =   element_blank(),        ## <- this line
        axis.text.x = element_blank()  )       ## <- and this line
        
Cases_Race
```

I pulled a subset of the defendant data with the variables I was interested in: docket_id, gender, race, and DOB.

```{r }
defSubset <- def %>% select(docket_id, gender, race, date_of_birth)

```

## Offenses Data

Overview of offense data

```{r }
glimpse(off)

```

I'm still wondering what the best way is to merge the offenses data with the bail data.

For this, I was primarily interested in the "grade" of the offense (i.e. first degree felony - F1). 
- H = murder (I confirmed this by spot checking records)
- F = felony
- M = misdemeanor

It's a bit challenging to look at the grade of the offense because some dockets have multiple offenses.  It's also difficult to rank the offenses by level of severity because some of them don't come with a degree  For example, I wouldn't know how to rank "F" vs "F3".  There's also some grades that weren't included in the ranking system we were provided, such as IC, S, S1, S2, or S3.  Also, sometimes there is no grade assigned.

We could also wait until the team who worked with the offense data shares their cleaned data files, and then after that we can merge bail with the offense data.

```{r }
#all the grades for offenses
unique(off$grade)
length(unique(off$grade)) #number of grade types

```

I decided to conduct a preliminary approach at getting the most severe grade/offense, since this one would likely have the biggest impact on bail. 

I put the grades in order as H, F, M, S, then IC.  I also put things order by degree, with grades without a degree last.

An issue with this approach is that a grade without a degree may not actually be less severe than a grade with a degree.  Also, I'm not sure what S and IC signify, so I'm not sure which would be more severe than the other.

```{r }
#attempting to order the factor levels by grade
off$grade = factor(off$grade, levels = c("H1", "H2", "F1", "F2", "F3", "F", "M1", "M2", "M3", "M", "S1", "S2", "S3", "S", "IC"), 
                   ordered = TRUE)

#reversed the leveling because I accidentally inputted the grades backwards
library(forcats)
off$grade = fct_rev(off$grade)

```

With just a preliminary look at the offense data, there's quite a few offenses with no grade. There's also quite a few offenses that have just a letter for the grade but no degree.

Once again, the team looking at the offense data will probably have more plots and insights.

```{r }
offGrade <- group_by(off, grade) %>% summarise(
  n_offenses = length(grade))

#plots a bar chart with the number of cases per offense grade
Cases_Grade <- ggplot(offGrade, aes(x=grade, y=n_offenses)) + 
  geom_bar(stat = "identity") + 
  labs(title="Number of Offenses per Grade",  
       y="Num Offenses", x="Offense Grade", subtitle ="Some cases may have mutliple offenses" )

Cases_Grade 

```

Creating a dataset grouped by docket id with the each case's most serious offense + number of offenses.

```{r }
#grouping the data, such that we will have one record per docket_id for offenses
offGrouped <- group_by(off, docket_id) %>% summarise(
  n_offenses = length(grade),
  most_ser_off = max(unique(grade)))
    
```

## Merging Bail with Defendent and Offense Data

Merging the grouped data from bail with the defendent data and then the grouped offenses data. This produces a dataset with one record per docket_id.

Upon merging, there's only missing values for most_ser_off. It makes sense that there are missing values for this variable because we saw that in the original offenses dataset; some cases did not have grades for offenses.

```{r }

joined <- left_join(bailGrouped, defSubset, by = "docket_id") %>%
          left_join(offGrouped, by = "docket_id")

#checking for missing data in the joined dataset 
plot_missing(joined)
    
```

## Considering "Before" and "After" 

Creating a subset of the data within 2016-2019 for two years before (2016-2017) and after (2016-2017) the new DA started. I also added dummy variables for before and after. 

```{r }

joined16_19 <- joined %>% filter(date_first_bail <= '2019-12-31') %>%
  filter(date_first_bail >= '2016-01-01')%>%
  mutate(before = ifelse(date_first_bail < '2018-01-01', 1, 0)) %>%
  mutate(after = ifelse(date_first_bail >= '2018-01-01', 1, 0))

```

## Bail by Race

Looking at trends in initial bail by race.

I focused on black and white because the other races were appeared infrequently relatively.

Data below considers for all offense types.

```{r }

ba_bw <- joined16_19 %>% filter(race %in% c("Black", "White")) %>% 
  mutate(race_time = interaction(race, after)) #adding a variable to show race and time point (before/after)
    
```

Creating a box plot of initial bail by race, considering before and after for all charge types.

scale_y_log10 was giving me issues because there are some bail values of zero. I added 1 to all the bail values to address this, but I'm wondering if there was a better solution.

```{r }

#added 1 to first bail so we can use the zero values even when we do a log transform 
ba_bw <- ba_bw %>% mutate(first_bail = first_bail + 1)

#box plot ofinitial bail by race - before and after - all offense types
race_bail <- ggplot(ba_bw, aes(x=race_time, y=first_bail, fill=after)) + 
  geom_boxplot(coef=1e12) + 
  labs(title="Initial Bail by Race + Before v After",  
       y="Initial Bail", x="Race.After", subtitle = "For all offense types")+
  scale_y_log10(limits=c(10^0, 10^7), breaks=10^(0:7))
race_bail
    
```

Data below filters for first degree felonies (F1) as the most serious offense grade. I chose to filter by F1 because it is the most severe grade after murder (and cases of murder occured infrequently). Based off the way I ranked offense grades, F1 is among the more likely to be accurately ranked in severity.  

```{r }

F1_ba_bw <- ba_bw %>%  filter(most_ser_off == "F1") 

#box plot ofinitial bail by race - before and after - offense grade F1 only
F1_race_bail <- ggplot(F1_ba_bw, aes(x=race_time, y=first_bail, fill=after)) + 
  geom_boxplot(coef=1e12) + 
  labs(title="Initial Bail by Race + Before v After",  
       y="Initial Bail", x="Race.After", subtitle = "For cases with first degree felonies")+
  scale_y_log10(limits=c(10^0, 10^7), breaks=10^(0:7))

F1_race_bail
    
```

Outstanding questions:

- How should we choose initial bail when there are two different values for bail in the first action date?
- If we choose to consider offense severity alongside bail, how should we deal with cases that have multiple offense grades?  
- When plotting initial bail vs race, how should I use scale_y_log10 when some bail amounts are zero?
- What does a zero value for bail actually mean?

Possible areas for future exploration

- Trends for gender and age
- Trends for judge actions (and how do we actually approach this?)

