---
title: "Explore bails"
output:
  html_document:
    theme: lumen
    toc: true
    code_download: true
    highlight: tango
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "htmls") })
---

# Team 3: Understanding Systemic Trends

Examine how things have changed in the courts since Larry Krasner took office in January 2018.
Some example trends of interest:

- Number of people charged overall and by type of charge
- Harshness in bail and/or sentencing
- Number of defendants who fail to pay low monetary bail
- Trends in case dismissals, probation, parole, etc.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library(lubridate)
library(ggradar)

colorBlindGrey6 <- c("#0072B2", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#CC79A7")

theme_set(theme_minimal())
theme_update(panel.grid.minor = element_blank())
```

Uncomment the following chunk to filter to the periods before and after Larry Krasner took office.
```{r}
# bail_short <- vroom::vroom('csv/bail.csv')
# # skim(bail_short)
# # str(bail_short)
# # table(bail_short$action_type_name)
# 
# larry_date <- as.Date('2018-01-31')
# start_period_1 <- larry_date - period(2, 'years')
# end_period_2 <- larry_date + period(2, 'years')
# 
# bail_larry <- bail_short %>%
#   mutate(period = case_when(
#     action_date < larry_date & action_date >= start_period_1 ~ 'before',
#     action_date >= larry_date & action_date < end_period_2 ~ 'after',
#     TRUE ~ NA_character_)) %>%
#   filter(!is.na(period))
# 
# save(bail_larry, file = 'processed-data/bail-larry.rds')
```

```{r}
load('processed-data/bail-larry.rds')
```


Changes in judges involvement:

```{r fig.width=10}
bail_judges <- bail_larry %>%
  mutate(judge_name = case_when(
    is.na(participant_name__first_name) ~ NA_character_,
    TRUE ~  paste(participant_name__first_name,
                            participant_name__last_name)))

judges <- bail_judges %>%
  count(period, docket_id, judge_name) %>% 
  drop_na(judge_name) %>% 
  group_by(docket_id) %>% 
  mutate(weighted_n = n/sum(n)) %>% 
  ungroup() %>% 
  group_by(period, judge_name) %>% 
  summarise(total_adj_n = sum(weighted_n)) 

top_judges <- judges %>% 
  top_n(10, total_adj_n) %>% 
  pull(judge_name)

judges %>% 
  filter(judge_name %in% top_judges) %>% 
  pivot_wider(names_from = judge_name, 
              values_from = total_adj_n, 
              values_fill = 0) %>% 
  ggradar(values.radar = c(NA, '2,500', '5,000'),
        group.colours = colorBlindGrey6,
        grid.mid = 2500,
        grid.max = 5000,
        base.size = 1,
        legend.position = "top") + 
  labs(title = 'Top judges involvement in bails') +
  guides(color = guide_legend(reverse = TRUE))
```

```{r}
min_bail <- bail_larry %>% 
  group_by(docket_id) %>% 
  mutate(min_amount = min(total_amount)) %>% 
  ungroup()

min_bail %>% 
  ggplot(aes(x = period, y = min_amount)) +
  geom_boxplot() +
  scale_y_log10()
```


```{r include=FALSE}
# Check docket_id 325683, 196349, 124026 or 150893 

# bail_larry %>% 
#   distinct(docket_id, total_amount) %>% 
#   count(docket_id) %>% 
#   filter(n > 2) %>% 
#   {.}

# bail_short %>% 
#   count(docket_id) %>% 
#   filter(n > 100)

# bail_short %>% 
#   filter(docket_id == 124026) %>% 
#   View()

# bail_larry %>% 
#   count(participant_name__title, period) %>% 
#   filter(!is.na(participant_name__title), n > 100) %>% 
#   ggplot(aes(x = n, 
#              fill = period, 
#              y = participant_name__title %>% fct_reorder(n))) +
#   geom_col(position = 'dodge2') +
#   scale_x_log10() +
#   labs(y = NULL, x = 'Frequency')

# min_bail %>% 
#   group_by(period) %>% 
#   count(docket_id) %>% 
#   ggplot(aes(x = period)) +
#   geom_bar() +
#   # scale_y_log10() +
#   NULL

# Percentage columns is largely uninformative.
# bail_larry$percentage %>% summary()

```
