---
title: "ROR"
author: "Katrina Gutierrez"
date: "4/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rates of ROR 

Rate of ROR: proportion of cases where defendants were released on recognizance and did not have to pay bail

Goals:

- Explore how rates of ROR have changed after Krasner took office
- Explore any racial disparities rates of ROR

Findings:

- Overall, more the rate of ROR increased after Krasney took office for both misdemeanors and felonies 
- For misdemeanors, racial disparities in receiving ROR increased between black and white defendants
- For felonies, rates of ROR seemed roughly equal for black and white defendants. 

Caveat: 

- We don't have hispanic/latino in our race data. It's possible this information is captured by Ethnicity, which we don't have.  

Krasner's No Cash Bail Policy

- Krasner established a policy where his office would not request bail for 25 types of offesnes in February, 2018, shortly after he took office.
- It's difficult to look closer at this policy with our data because there are certain details we don't have.  For example, possession with intent deliver made up almost a third of eligible cases.  However, these cases are inelible if the a defendant committed the same offense within six months prior.  We don't have to distinguish these cases.  
- The exploration in this file considers cases regardless of their eligibility for the no cash bail policy.

## Initial set-up

Most of the initial set up is copied and pasted from "bail and race"
I used the code from that file as a basis to create the graphs in this file.

Loading libraries

```{r }
library(tidyverse)
library(lubridate)
library(DataExplorer)
library(vroom)
library(stringr)

theme_set(theme_minimal())
theme_update(panel.grid.minor = element_blank())
```

Loading the full datasets (all years):

```{r}
larry_date <- as.Date("2018-01-02")
start_period_1 <- larry_date - period(2, "years")
end_period_2 <- larry_date + period(2, "years")
dat_dir <- "csv"

bail_raw <- vroom::vroom(file.path(dat_dir, "bail.csv"))
off_raw <- vroom::vroom(file.path(dat_dir, "offenses_dispositions.csv")) %>%
  mutate(
    grade =
      grade %>% # forcats will default to alpha order
        fct_relevel("H1", "H2", after = 0) %>% # move these to the front
        fct_relevel("IC", after = Inf) %>% # move this to the back
        fct_rev() %>%
        factor(ordered = TRUE)
  )
def_raw <- vroom::vroom(file.path(dat_dir, "defendant_docket_details.csv")) %>%
  replace_na(list(race = "Unknown/Unreported"))
```

Initial cleaning for bail data

```{r}
bail <- bail_raw %>%
  # releveling action_type_name so that "Set" comes first
  mutate(action_type_name = action_type_name %>%
    as.factor() %>%
    relevel(ref = "Set")) %>%
  filter(
    participant_name__title == "Judge",
    !is.na(participant_name__first_name)
  ) %>%
  mutate(
    judge = paste(
      participant_name__last_name,
      participant_name__first_name,
      sep = ","
    ),
    action_date = ymd(action_date),
    action_type_name = factor(action_type_name)
  )
```


Creating a dataset where each record is a docket_id with summary bail info 

```{r cache=TRUE}
bail_grouped <- bail %>%
  group_by(docket_id) %>%
  summarise(
    n_judges = length(unique(judge)),
    min_bail = min(total_amount),
    max_bail = max(total_amount),
    date_first_bail = first(action_date),
    first_bail = first(total_amount),
    first_bail_type = factor(first(type_name)),
    ROR_ever = ifelse(any(type_name == "ROR"), 1, 0) # did case ever receive ROR
  )
```

Creating a subset of the defendant data 

```{r }
def_subset <- def_raw %>% select(docket_id, gender, race, date_of_birth)
```

Creating a dataset grouped by docket id with the each case's most serious offense + number of offenses.

```{r cache=TRUE}
offense_grouped <- off_raw %>%
  filter(!is.na(grade)) %>%
  group_by(docket_id) %>%
  summarise(
    n_offenses = length(grade),
    most_ser_off = max(unique(grade))
  )

```

Merging the grouped data from bail with the defendent data and then the grouped offenses data.
This produces a dataset with one record per docket_id.

```{r }
joined <- bail_grouped %>%
  left_join(def_subset, by = "docket_id") %>%
  left_join(offense_grouped, by = "docket_id") %>%
  filter(!is.na(ROR_ever)) %>%
  mutate(period = case_when(
    between(date_first_bail, start_period_1, larry_date) ~ "before",
    between(date_first_bail, larry_date, end_period_2) ~ "after",
    TRUE ~ NA_character_
  )) %>%
  mutate(grade = ifelse(most_ser_off %in% c("F1", "F2", "F3", "F"), "felony",
    ifelse(most_ser_off %in% c("M1", "M2", "M3", "M"), "misdemeanor", NA)
  ))



joined$period <- relevel(factor(joined$period), ref = "before")
```

## Rate of ROR: Misdemeanors vs Felonies

```{r }
all <- joined %>%
  filter(!is.na(period)) %>%
  filter(!is.na(grade)) %>%
  group_by(period, grade) %>%
  summarise(
    n_dockets = length(ROR_ever),
    n_ROR = sum(ROR_ever),
    ROR_rate = n_ROR / n_dockets,
  )
all %>%
  ggplot(aes(
    x = grade,
    y = ROR_rate,
    fill = period
  )) +
  geom_col(position = "dodge") +
  labs(
    title = "Rate of ROR",
    y = "Proportion of Cases with ROR", x = "Grade"
  ) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_cartesian(expand = FALSE)
```

## Misdemeanors: ROR + Race


```{r }
misdemeanor <- joined %>%
  filter(grade == "misdemeanor")
```

ROR rates for Misdemeanors by race before and after.  White people disproportionally benefitted 

Looked at black and white because these were the most common races.

```{r }

misdemeanor3 <- misdemeanor %>%
  filter(!is.na(period)) %>%
  filter(race %in% c("Black", "White")) %>%
  group_by(period, race) %>%
  summarise(
    n_dockets = length(ROR_ever),
    n_ROR = sum(ROR_ever),
    ROR_rate = n_ROR / n_dockets,
  )

misdemeanor3 %>%
  ggplot(aes(
    x = race,
    y = ROR_rate,
    fill = period
  )) +
  geom_col(position = "dodge") +
  labs(
    title = "Rate of ROR - Misdemeanors",
    y = "Proportion of Cases with ROR", x = "Race"
  ) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_cartesian(expand = FALSE)
```

## Felonies: ROR + Race

ROR rates for felonies in general for all races went up from 8 to 13% after Larry

```{r }

felony <- joined %>%
  filter(grade == "felony") %>%
  filter(!is.na(period))


# ROR rates for felonies black vs white before and after

felony3 <- felony %>%
  filter(!is.na(period)) %>%
  filter(race %in% c("Black", "White")) %>%
  group_by(period, race) %>%
  summarise(
    n_dockets = length(ROR_ever),
    n_ROR = sum(ROR_ever),
    ROR_rate = n_ROR / n_dockets,
  )

felony3 %>%
  ggplot(aes(
    x = race,
    y = ROR_rate,
    fill = period
  )) +
  geom_col(position = "dodge") +
  labs(
    title = "Rate of ROR - Felonies",
    y = "Proportion of Cases with ROR", x = "Race"
  ) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_cartesian(expand = FALSE)
```
